---
title: "transcripts-LPS-code"
author: "K. Martinez, K. Wilding"
date: "2025-01-13"
output: 
  html_document: 
    fig_height: 10
    code_folding: show
    fig_width: 15
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 6
    highlight: default
editor_options: 
  chunk_output_type: console
---

# Copyright
© 2025. Triad National Security, LLC. All rights reserved.

This program was produced under U.S. Government contract 89233218CNA000001 for Los Alamos National Laboratory (LANL), which is operated by Triad National Security, LLC for the U.S. Department of Energy/National Nuclear Security Administration. All rights in the program are reserved by Triad National Security, LLC, and the U.S. Department of Energy/National Nuclear Security Administration. The Government is granted for itself and others acting on its behalf a nonexclusive, paid-up, irrevocable worldwide license in this material to reproduce, prepare. derivative works, distribute copies to the public, perform publicly and display publicly, and to permit others to do so.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up


## Environment Set Up

```{r Libraries and Directories, echo=TRUE, message=FALSE, warning=FALSE}
rm(list=ls())
working.dr <- "PATH"
setwd(working.dr)
# Model Libraries
library(randomForest)
library(glmnet)
library(naivebayes)
library(neuralnet)
library(nnet)
library(e1071)
library(klaR)
library(kernlab)
library(RSNNS)
## Basic Libraries
library(plyr)
library(tidyverse)
library(xtable)
library(readxl)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(colorspace)
library(caret)
library(ggpubr)
library(pROC)


```

```{r Function Definitions}
multiplesheets <- function(fname,sheet_nums) {
  # getting info about all excel sheets
  sheets <- readxl::excel_sheets(fpath)[sheet_nums]
  ## make list from sheets determined by sheet_nums. Skip 2 rows, and repair names
  tibble <- lapply(sheets, function(x) readxl::read_excel(fpath, sheet = x, skip=2,.name_repair = make.names))
  data_frame <- lapply(tibble, as.data.frame)
  # assigning names to data frames
  names(data_frame) <- make.names(sheets)
  # print data frame
  return(data_frame)
}
```

# Data Import, Clean, Normalization

The primary structure of this code and background on the data comes from LPS_qPCR_Analysis_ModelBuild.Rmd. Refer to that file for more details. This is just getting all the data into a tidy format and then

### Raw Data Sets

#### Primary LPS Data

```{r Import Data, message=FALSE, warning=FALSE}
# specifying the path name
fpath <- "data/Int_Imm_Sample_Spreadsheet_current.xlsx"
LPS.meta.data <- read_excel(fpath,sheet=excel_sheets(fpath)[1],.name_repair = make.names)
LPS.qPCR.list.plates <- multiplesheets(fpath,2:length(excel_sheets(fpath)))
```

```{r Clean and Organize Data, message=FALSE, warning=FALSE}
LPS.qPCR.tidy <-  LPS.qPCR.list.plates %>% 
  ## map() allows us to apply these dplyr fxns across all dfs in the list
  map(~.x %>% 
        mutate(across(starts_with("X"), ~ as.numeric(replace(.x,.x=="Undetermined",41)))) %>% 
        pivot_longer(starts_with("X"),names_to="SampleIndex",values_to="CT_value")
  ) %>% 
  bind_rows() %>% 
  filter(!is.na(CT_value)) %>% ## filter out experiments not run yet
  mutate(SampleIndex=gsub("X","",SampleIndex)) %>% 
  left_join(LPS.meta.data, 
            by=c("SampleIndex"="X3.Letter.Code"))

# LPS.qPCR.wide <-  bind_rows(LPS.qPCR.list.tidy) %>% pivot_wider(names_from = SampleIndex,values_from=CT_value)
```

#### Delayed/New Tech LPS Exp

```{r MJ LPS Data, message=FALSE, warning=FALSE}
fpath <- "data/09012022_ctvalues_noLPSvsLPS_bothexperiments.xlsx"
MJ.LPS.qPCR.raw <- read_excel(fpath)[-1,-1] 
MJ.LPS.qPCR.tidy <- MJ.LPS.qPCR.raw %>% 
  pivot_longer(!Symbol,names_to="SampleIndex",values_to="CT_value") %>% 
  mutate(LPS.Added=ifelse(grepl("no",SampleIndex),0,1)) %>% 
  mutate(CT_value=as.numeric(ifelse(CT_value=="Undetermined","41",CT_value))) %>% 
  rename(Gene.Symbol=Symbol)
```

#### LPS Smaller Panel

## Normalization to HK: B2M

This is the normalization to only B2M

```{r Normalization 1, message=FALSE, warning=FALSE}
LPS.qPCR.norm1.tidy <- full_join(LPS.qPCR.tidy %>%
                                   filter(Gene.Symbol%in%c("B2M")) %>%  
                                   group_by(SampleIndex) %>% 
                                   summarise(HK_mean=mean(CT_value,na.rm = T)),
                                 LPS.qPCR.tidy %>% 
                                   mutate(CT_value=ifelse(CT_value==41,NA,CT_value))) %>% 
  mutate(CT_value_norm1=CT_value-HK_mean) %>% 
  mutate(CT_value_norm1=ifelse(is.na(CT_value_norm1),max(CT_value_norm1,na.rm = T)+1,CT_value_norm1))
#MJ LPS
MJ.LPS.qPCR.norm1.tidy <- full_join(MJ.LPS.qPCR.tidy %>%
                                     filter(Gene.Symbol%in%c("B2M")) %>%  
                                     group_by(SampleIndex) %>% 
                                     summarise(HK_mean=mean(CT_value,na.rm = T)),
                                    MJ.LPS.qPCR.tidy %>% 
                                     mutate(CT_value=ifelse(CT_value==41,NA,CT_value))) %>% 
  mutate(CT_value_norm1=CT_value-HK_mean) %>% 
  mutate(CT_value_norm1=ifelse(is.na(CT_value_norm1),max(CT_value_norm1,na.rm = T)+1,CT_value_norm1)) 
```

## Model Data Format

```{r All Model Data, message=FALSE, warning=FALSE}
# Primary LPS
LPS.model.data <- LPS.qPCR.norm1.tidy %>% 
  filter(!(Gene.Symbol%in%c("ACTB","B2M","GAPDH","HPRT1","RPLP0","HGDC","RTC","PPC"))) %>%  ## Filter out Controls
  select(LPS.Added,CT_value_norm1,Gene.Symbol,SampleIndex) %>%
  group_by(SampleIndex) %>% 
  ungroup() %>% 
  pivot_wider(names_from=Gene.Symbol,values_from=CT_value_norm1) %>% 
  select(-SampleIndex) %>%
  mutate(LPS.Added=ifelse(LPS.Added=="Y",1,0)) %>% 
  rename(IL13="1L13") ## I think this is supposed to be IL13 not 1L13
#MJ LPS 
MJ.LPS.model.data <- MJ.LPS.qPCR.norm1.tidy %>% 
  filter(!(Gene.Symbol%in%c("ACTB","B2M","GAPDH","HPRT1","RPLP0","HGDC","RTC","PPC"))) %>%  ## Filter out Controls
  select(LPS.Added,CT_value_norm1,Gene.Symbol,SampleIndex) %>%
  group_by(SampleIndex) %>% 
  ungroup() %>% 
  pivot_wider(names_from=Gene.Symbol,values_from=CT_value_norm1) %>% 
  # select(-Proto_SampleIndex) %>%
  mutate(LPS.Added=ifelse(LPS.Added==1,1,0)) #%>% 
  # rename(IL13="1L13")

```

# Train/Test Split Stability

This will run a train/test split stability analysis for all the considered models at the same time. The models used will use the default hyper-parameters where possible to minimize structural differences among trained models.

-   Random Forest: mtry - sqrt(84)=9.1615, ntree=500

-   Logistic Regression: lambda =  .01 (based on a cv.glmnet)

-   SVM: kernel=linear, Cost = 1

-   Naive Bayes: usekernel=TRUE (non-parametric kernel density - no normal assumption)

-   Single Layer Neural Net: size =??? , decay = 0, maxit=100 size determined by preliminary tuning(see directly below)

-  Do not need this anymore, as RFE only supports single layer nnet* 2 hidden layer NN: size 42, 21, error function= "ce", linear output = FALSE* 

Clearly there are many more options for the hyper parameters, but as this is a test to see what a good train/test split is for limited data, simple hyper parameter tunning will have to come next. \## Functions For Train/Test Split

```{r NN basic tune size}
# trainctrl <- trainControl(method="repeatedcv", number=10, repeats=5)
# nnetGrid <-  expand.grid(size = seq(from = 1, to = 11, by = 1),
#                         decay = c(0,.5,.1,.05,.01,.001,.0001))
# nn.pre.mod <- train(x=LPS.model.data[,-1],as.factor(LPS.model.data$LPS.Added),
#                     method="nnet",tuneGrid = nnetGrid,trace=FALSE,linout=FALSE,trControl = trainctrl)
```

```{r RF basic tune size}
# trainctrl <- trainControl(method="repeatedcv", number=10, repeats=2)
# rfGrid <-  expand.grid(mtry = seq(from = 1, to = 10, by = 1))
# rf.pre.mod <- train(x=LPS.model.data[,-1],as.factor(LPS.model.data$LPS.Added),
#                     method="rf",tuneGrid = rfGrid,trControl = trainctrl)
# any value between 2 and 39 is justifiable. So using default sqrt(84) makes sense.
```

```{r GLM basic tune}
# glm.pre.mod <- cv.glmnet(x=as.matrix(LPS.model.data[,
#                                           -which(colnames(LPS.model.data) ==
#                                                      "LPS.Added")]),
#                    y = as.vector(LPS.model.data$LPS.Added),family = "binomial",type.measure = "class",
#                    lambda=c(5e-1,1e-1,5e-2,1e-2,5e-3,1e-3,5e-4,1e-4,5e-5,1e-5))
# glm_grid=expand.grid(lambda=c(1e-1,1e-2,1e-3,1e-4,1e-5),
#                      alpha=1)
# glm.pre.mod <- train(x=LPS.model.data[,-c(1)],y=as.factor(LPS.model.data$LPS.Added),
#                    method="glmnet",family="binomial",tuneGrid=glm_grid)


### Everytime this is run you get WILDLY different answers. ranging from .01-.1 picking one that makes sense.
```

```{r SVM basic tune}
# svm.pre.mod <- tune(svm,as.factor(LPS.Added)~.,data=LPS.model.data,kernel='linear',ranges=list(cost = c(1,1e-1,1e-2,1e-3,1e-4,1e-5)))
# svm.pre.mod$best.parameters
### Everytime this is run you get WILDLY different answers. ranging from .005-5 picking one that makes sense.

```

```{r General Error Function, message=FALSE, warning=FALSE}
generate.errors.ML <- function(model.string,model.obj,predictors){
  y <- predictors$LPS.Added
  if(model.string %in% c("RF","SVM")){
    observed <- as.factor(y)
    CM <- caret::confusionMatrix(predict(model.obj, predictors),observed)
  } else if(model.string=="NB"){
    observed <- as.factor(as.logical(y))  
    CM <- caret::confusionMatrix(predict(model.obj, predictors),observed)
    } else if(grepl("NN3",model.string)){
      observed <- as.factor(y)
      predictorsnn3 <- predictors[,-1]
      CM <- caret::confusionMatrix(factor(round(predict(model.obj,predictorsnn3)[,1]),levels=c(0,1)),observed)
    } else if(grepl("NN",model.string)){
      observed <- as.factor(y)
      CM <- caret::confusionMatrix(factor(predict(model.obj,newdata=predictors,type="class"),levels=c(0,1)),observed)
    } else if(model.string=="GLM"){
      # observed <- as.vector(y)
      # CM <- caret::confusionMatrix(
      #   as.factor(predict(model.obj,
      #                     as.matrix(predictors[, -which(colnames(predictors) ==
      #                                                     "LPS.Added")]),type="class",s='lambda.min')),
      #   as.factor(observed))
      observed <- as.factor(y)
      CM <- caret::confusionMatrix(as.factor(predict(model.obj,newx = as.matrix(predictors[,-1]),type="class")),
                            observed)
    } 
  error.vec <- c(CM$overall,CM$byClass)
  return(error.vec)
}

## Model Errors
    error_tab <- function(model.string,model.obj,train_data,test_data){
      error_train <- generate.errors.ML(model.string,model.obj,predictors=train_data) %>% 
        data.frame(error_val=.) %>% 
        rownames_to_column(var="error_type") %>% 
        mutate(split_data="train")
      error_test <- generate.errors.ML(model.string,model.obj,predictors=test_data)%>% 
        data.frame(error_val=.) %>% rownames_to_column(var="error_type") %>% 
        mutate(split_data="test")
      error_mat <- bind_rows(error_train,error_test) %>% mutate(model_type=model.string)
      return(error_mat)
    }

```

```{r Train/Test Split Robustness Function}
# # For Testing
# model.data <- LPS.model.data
# nrep <- 2

train_test_stability <- function(nrep=5,model.data){
  split.vec <- seq(0,1,by=.02)
  split.vec <- split.vec[-c(1:5,(length(split.vec)-4):length(split.vec))]
  # split.vec <- split.vec[c(1,5,10,15)]
all.models <- vector("list",length(split.vec))
names(all.models) <- split.vec
for(i in 1:length(split.vec)){
  prop=split.vec[i]
  model.group <- vector("list", nrep)
  names(model.group) <- 1:nrep
  message(paste("Starting",split.vec[i]))
  for(k in 1:nrep){
    model.obj <- vector("list",4)
    names(model.obj) <- c("split.index",
                          "train.scale",
                          "models",
                          "errors") 
    # Train Test Split
    train_index <- createDataPartition(model.data$LPS.Added,p=prop,list=F)
    model.obj[["split.index"]] <- train_index
    # train_data <- model.data[train_index,]
    # test_data <- model.data[-train_index,]
    
    # Scale Data to training data
    train.min.max <- model.data[train_index,] %>%
      pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
      group_by(Gene) %>%
      summarise(CTmin=min(CT_norm1),
                CTmax=max(CT_norm1)) %>%
      ungroup()
    model.obj[["train.scale"]] <- train.min.max
    
    train_data <- model.data[train_index,] %>%
      pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
      full_join(train.min.max,by="Gene") %>%
      mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
      mutate(CT_scaled=ifelse(is.finite(CT_scaled),CT_scaled,1)) %>% 
      select(LPS.Added,Gene,CT_scaled) %>%
      group_by(Gene) %>%
      mutate(row_num=row_number()) %>%
      pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
      select(-row_num)
    
    test_data <- model.data[-train_index,] %>%
      pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
      full_join(train.min.max,by="Gene") %>%
      mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
      mutate(CT_scaled=ifelse(is.finite(CT_scaled),CT_scaled,1))%>%
      select(LPS.Added,Gene,CT_scaled) %>%
      group_by(Gene) %>%
      mutate(row_num=row_number()) %>%
      pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
      select(-row_num)
    # Train Models 
    # function(model.string,model.obj,predictors)
    mods.list <- vector("list", 6)
    names(mods.list) <- c("rf","nb","svm","glm","nn","nn3")
    
    mod.rf <- randomForest(as.factor(LPS.Added)~., data=train_data,
                           proximity=TRUE,importance=TRUE,ntree=500,keep.forest=TRUE)
    mods.list[["rf"]] <- mod.rf
    
    mod.nb <- naive_bayes(as.logical(LPS.Added) ~ ., data = train_data, usekernel = T, type="raw")
    mods.list[["nb"]] <- mod.nb
    
    mod.svm <- svm(as.factor(LPS.Added)~.,data=train_data,cost=.1, kernel='linear')
    mods.list[["svm"]] <- mod.svm
    
    # mod.glm <- cv.glmnet(x = as.matrix(train_data[, 
    #                                               -which(colnames(train_data) == 
    #                                                        "LPS.Added")]), 
    #                      y = as.vector(train_data$LPS.Added), 
    #                      type.measure = "class", 
    #                      family = "binomial")
    mod.glm <- glmnet(x=train_data[,-1],
                   y=as.factor(train_data$LPS.Added),
                   family = "binomial",lambda = .01)
    mods.list[["glm"]] <- mod.glm

    mod.nn <- nnet(as.factor(LPS.Added) ~ ., data = train_data ,size=8,linout = FALSE,trace=F,decay=.1)
    mods.list[["nn"]] <- mod.nn
    
    # library(RSNNS)
    mod.nn3 <- mlp(x=train_data[,-1],
                   y=train_data$LPS.Added , size = c(80,20,29))
    mods.list[["nn3"]] <- mod.nn3

    
    model.obj[["models"]] <- mods.list
    
    ## Model Errors
    error_tab <- function(model.string,model.obj,train_data,test_data){
      error_train <- generate.errors.ML(model.string,model.obj,predictors=train_data) %>% 
        data.frame(error_val=.) %>% 
        rownames_to_column(var="error_type") %>% 
        mutate(split_data="train")
      error_test <- generate.errors.ML(model.string,model.obj,predictors=test_data)%>% 
        data.frame(error_val=.) %>% rownames_to_column(var="error_type") %>% 
        mutate(split_data="test")
      error_mat <- bind_rows(error_train,error_test) %>% mutate(model_type=model.string)
      return(error_mat)
    }
    errors <- bind_rows(error_tab(model.string="RF",model.obj=mod.rf,train_data,test_data),
                        error_tab(model.string="GLM",model.obj=mod.glm,train_data,test_data),
                        error_tab(model.string="SVM",model.obj=mod.svm,train_data,test_data),
                        error_tab(model.string="NN",model.obj=mod.nn,train_data,test_data),
                        error_tab(model.string="NN3",model.obj=mod.nn3,train_data,test_data),
                        error_tab(model.string="NB",model.obj=mod.nb,train_data,test_data)) %>% 
      filter(error_type%in%c("Accuracy","Sensitivity","Specificity","Precision"))
    model.obj[["errors"]] <- errors
    # auc <- bind_rows(error_tab(model.string="RF",model.obj=mod.rf,train_data,test_data),
    #                     error_tab(model.string="GLM",model.obj=mod.glm,train_data,test_data),
    #                     error_tab(model.string="SVM",model.obj=mod.svm,train_data,test_data),
    #                     error_tab(model.string="NN",model.obj=mod.nn,train_data,test_data),
    #                     error_tab(model.string="NN3",model.obj=mod.nn3,train_data,test_data),
    #                     error_tab(model.string="NB",model.obj=mod.nb,train_data,test_data))
    # model.obj[["auc"]]
    # mod.nn
    # mod.glm
    # 
    # Model Errors
    # model.obj[["model"]] <- mod
    model.group[[k]] <- model.obj
    if(k%%100==0){
      message(paste("Progress",split.vec[i],":",k))
    }
  }
  all.models[[i]] <- model.group
  # message(paste("Finished",split.vec[i]))

}
error.data.frame <- data.frame()

for(sp in 1:length(all.models)){
  error.data.frame <- bind_rows(error.data.frame,bind_rows(purrr::map(all.models[[sp]],~.x[["errors"]] %>%
                                             mutate(SplitPerc=as.numeric(names(all.models)[sp]))),
                                .id = "rep"))
}
# save(all.models,error.data.frame,model.data,file = paste0("output/traintest_split_stability_",nrep,".RData"))
return(list(all.models,error.data.frame))
}
run_yn=T
save_yn=T
if(run_yn==T){
  system.time({
  output <- train_test_stability(nrep=200,model.data = LPS.model.data)
  })
  error.data.frame <- output[[2]]
  if(save_yn==T){
    write.csv(error.data.frame, 
              paste0(working.dr,"data/traintest_split_stability_errordf.csv"),row.names = FALSE)
  }
}
```

```{r Train/Test Split Analysis}

# load(paste0("output/traintest_split_stability_",nrep,".RData"))
files <- list.files(working.dr,pattern ="traintest_split",full.names = T)
files <- files[which(grepl(".csv",files))]
error.data.frame <- data.frame()
for(i in files){
  error.data.frame <- bind_rows(error.data.frame,read.csv(i))
}
ggplot(error.data.frame %>% 
         # filter(error_type=="Accuracy") %>% 
         filter(split_data=="test") %>% 
         group_by(SplitPerc,model_type) %>% 
         mutate(mean_error=mean(error_val,na.rm=T),
                sd_error=sd(error_val,na.rm=T),
                mean_sd_top=min(1,mean_error+sd_error),
                mean_sd_low=mean_error-sd_error) %>% 
         ungroup() %>% 
         filter(SplitPerc>=.1&SplitPerc<=.95),
       # aes(x=SplitPerc,y=1-error_val))+
       aes(x=SplitPerc,y=error_val))+
  geom_point(alpha=.01,size=1)+
  # geom_line(aes(y=1-mean_error),color="forestgreen",size=2)+
  geom_line(aes(y=mean_error),color="forestgreen",size=1.5)+
  # geom_line(aes(y=1-mean_sd_top),color="forestgreen",size=1.5,linetype=3,alpha=.9)+
  # geom_line(aes(y=1-mean_sd_low),color="forestgreen",size=1.5,linetype=3,alpha=.9)+
  # geom_ribbon(aes(ymax = 1-mean_sd_low, ymin = 1-mean_sd_top),linetype=2, alpha = 0.25)+
  geom_ribbon(aes(ymin = mean_sd_low, ymax = mean_sd_top),linetype=2, alpha = 0.25)+
  xlab("Training Data Percetage")+
  ylab("Testing Error")+
  facet_grid(error_type~model_type)+
  theme_clean()+ 
  scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        )+
  # coord_cartesian(ylim=c(0,.5),xlim=c(.1,.9))
  coord_cartesian(ylim=c(.5,1),xlim=c(.1,.9))


ggplot(error.data.frame %>% 
         filter(error_type=="Accuracy") %>% 
         filter(split_data=="test") %>% 
         group_by(SplitPerc,model_type) %>% 
         mutate(mean_error=mean(error_val,na.rm=T),
                sd_error=sd(error_val,na.rm=T),
                mean_sd_top=min(1,mean_error+sd_error),
                mean_sd_low=mean_error-sd_error) %>% 
         ungroup() %>% 
         filter(SplitPerc>=.1&SplitPerc<=.95),
       # aes(x=SplitPerc,y=1-error_val))+
       aes(x=SplitPerc))+
  # geom_point(alpha=.01,size=1)+
  # geom_line(aes(y=1-mean_error),color="forestgreen",size=2)+
  geom_line(aes(y=sd_error),color="forestgreen",size=1.5)+
  # geom_line(aes(y=1-mean_sd_top),color="forestgreen",size=1.5,linetype=3,alpha=.9)+
  # geom_line(aes(y=1-mean_sd_low),color="forestgreen",size=1.5,linetype=3,alpha=.9)+
  # geom_ribbon(aes(ymax = 1-mean_sd_low, ymin = 1-mean_sd_top),linetype=2, alpha = 0.25)+
  # geom_ribbon(aes(ymin = mean_sd_low, ymax = mean_sd_top),linetype=2, alpha = 0.25)+
  xlab("Training Data Percetage")+
  ylab("Testing SD")+
  facet_grid(.~model_type)+
  theme_clean()+ 
  scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        )#+
  # coord_cartesian(ylim=c(0,.5),xlim=c(.1,.9))
  # coord_cartesian(ylim=c(.5,1),xlim=c(.1,.9))

error.data.frame %>% 
  filter(split_data=="test") %>% 
  filter(error_type=="Accuracy") %>% 
  group_by(SplitPerc,model_type) %>% 
  summarise(mean_error=mean(error_val,na.rm=TRUE)) %>% 
  filter(mean_error==1) %>%
  arrange(model_type,SplitPerc)
library(ggokabeito)
error.data.frame %>% filter(error_type=="Accuracy") %>% 
  filter(split_data=="test") %>% 
  group_by(SplitPerc,model_type) %>% 
  count(error_val) %>% 
  mutate(Not100Perc=ifelse(error_val<1,TRUE,FALSE)) %>% 
  group_by(SplitPerc,model_type,Not100Perc,group=Not100Perc) %>% 
  summarise(total_mods=sum(n)) %>% 
  mutate(prop=total_mods/300) %>% 
  filter(!Not100Perc) %>% 
  ggplot(aes(x=SplitPerc,y=prop,color=model_type))+
  geom_line(size=2)+facet_wrap(.~Not100Perc)+
  ylab("Proportion of Models")+
  theme_clean()+ 
  scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        )+
  scale_color_okabe_ito()


test <- error.data.frame %>% filter(error_type=="Accuracy") %>% 
  filter(split_data=="test") %>% 
  group_by(SplitPerc,model_type) %>% 
  count(error_val) %>% 
  mutate(Not100Perc=ifelse(error_val<1,TRUE,FALSE)) %>% 
  group_by(SplitPerc,model_type,Not100Perc,group=Not100Perc) %>% 
  summarise(total_mods=sum(n)) %>% 
  mutate(prop=total_mods/500) %>% 
  filter(model_type=="NB") 
#%>% filter(prop>=.5&!Not100Perc) %>% group_by(model_type) %>% slice_min(SplitPerc)
#test %>% filter(SplitPerc==.7&!Not100Perc)
error.data.frame %>% filter(error_type=="Accuracy") %>% 
  filter(split_data=="test") %>% 
  group_by(SplitPerc,model_type) %>% 
  count(error_val) %>% 
  mutate(Not100Perc=ifelse(error_val<1,TRUE,FALSE)) %>% 
  group_by(SplitPerc,model_type,Not100Perc,group=Not100Perc) %>% 
  summarise(total_mods=sum(n)) %>% 
  mutate(prop=total_mods/500) %>% 
  filter(SplitPerc==.7) %>% 
  filter(!Not100Perc)

error.data.frame %>% filter(error_type=="Accuracy") %>% 
  filter(split_data=="test") %>% 
  group_by(SplitPerc,model_type) %>% 
  count(error_val) %>% 
  mutate(Not100Perc=ifelse(error_val<=.9,TRUE,FALSE)) %>% 
  group_by(SplitPerc,model_type,Not100Perc,group=Not100Perc) %>% 
  summarise(total_mods=sum(n)) %>% 
  mutate(prop=total_mods/500) %>% 
  filter(SplitPerc==.7) %>% 
  filter(!Not100Perc)

error.data.frame %>% 
         filter(error_type=="Accuracy") %>% 
         filter(split_data=="test") %>% 
         group_by(SplitPerc,model_type) %>% 
         mutate(mean_error=mean(error_val,na.rm=T)) %>% 
         ungroup() %>% 
         filter(SplitPerc==.7) %>% 
  select(model_type,mean_error,SplitPerc) %>% 
  distinct()
```

```{r Training Data}
set.seed(1306) # Finalzize seed for this command with a data set. These two lines need to be run back to back for consistent sample. 
## 4765 RF 1/.95/.833 (on Min/max Gene/Ind scaled)
## 1885 RF 1/.95/1 (on Min/max Gene/Ind scaled)
## 1306 RF 1/.95/1 (on Min/max Gene/Ind scaled)
## 444 RF 1/1/1 
train_index <- createDataPartition(LPS.model.data$LPS.Added,p=.7,list=F) 
# Once settled on training set save index and verify that there is no difference between what is generated above and the saved
train_index_saved <- as.matrix(read.csv("data/II_ML_LPS_train_index.csv"))
if(length(setdiff(train_index,train_index_saved))==0){
  train_data <- LPS.model.data[train_index,]
  test_data <- LPS.model.data[-train_index,]
  print("Match confirmed. Data Partitioned")
  
}else{
  print("Saved index and generated index DO NOT MATCH. Using saved list.")
  train_data <- LPS.model.data[train_index_saved,]
  test_data <- LPS.model.data[-train_index_saved,]
}
train.min.max <- train_data %>%
  pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
  group_by(Gene) %>%
  summarise(CTmin=min(CT_norm1),
            CTmax=max(CT_norm1)) %>%
  ungroup()

train_data_scaled <- train_data %>%
  pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
  full_join(train.min.max) %>%
  mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
  select(LPS.Added,Gene,CT_scaled) %>%
  group_by(Gene) %>%
  mutate(row_num=row_number()) %>%
  pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
  select(-row_num)

test_data_scaled <- test_data %>%
  pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
  full_join(train.min.max) %>%
  mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
  select(LPS.Added,Gene,CT_scaled) %>%
  group_by(Gene) %>%
  mutate(row_num=row_number()) %>%
  pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
  select(-row_num)

MJ.LPS_data_scaled <- MJ.LPS.model.data %>%
  pivot_longer(!(LPS.Added:SampleIndex),names_to = "Gene",values_to = "CT_norm1") %>%
  full_join(train.min.max) %>%
  mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
  select(LPS.Added,Gene,CT_scaled) %>%
  group_by(Gene) %>%
  mutate(row_num=row_number()) %>%
  pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
  select(-row_num)
```

```{r TT take 2}
generate.errors.ML <- function(model.string,model.obj,predictors){
  y <- predictors$LPS.Added
  if(model.string %in% c("RF")){
    observed <- as.factor(y)
    class_pred <- as.factor(predict(model.obj,predictors,type="response"))
    prob_pred <- predict(model.obj,predictors,type="prob")[,2]
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(model.string%in%c("SVM")){
    observed <- as.factor(y)  
    class_pred <- as.factor(predict(model.obj,predictors,type="response"))
    prob_pred <- predict(model.obj,predictors,type="probabilities")[,2]
    colnames(prob_pred) <- NULL
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(grepl("GLM",model.string)){
    observed <- as.factor(y)
    class_pred <- as.factor(predict(model.obj,newx = as.matrix(predictors[,-1]),type="class"))
    prob_pred <- predict(model.obj,as.matrix(predictors[-c(1)]),type="response")
    CM <- caret::confusionMatrix(class_pred,observed)
    # predict(cv_glm,newx = as.matrix(train_data_scaled[,-c(1)]),s=.44,type="response") # Probabilities(of 1)
    # predict(cv_glm,newx = as.matrix(train_data_scaled[,-c(1)]),s=.44,type="class") # class

  } else if(grepl("NB",model.string)){
    observed <- as.factor(y)
    class_pred <- as.factor(as.numeric(predict(model.obj,predictors,type="class"))-1)
    prob_pred <- predict(model.obj,predictors,type="prob")[,2]
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(grepl("NN3",model.string)){
    observed <- as.factor(y)  
    prob_pred <- predict(model.obj,as.matrix(predictors[-c(1)]))[,1]
    class_pred <- factor(round(predict(model.obj,predictors[-c(1)])[,1]))
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(grepl("NN",model.string)){
    observed <- as.factor(y)  
    class_pred <- as.factor(predict(model.obj,predictors,type="class"))
    prob_pred <- predict(model.obj,predictors,type="raw")[,1]
    CM <- caret::confusionMatrix(class_pred,observed)
  }
  error.vec <- c(CM$overall,CM$byClass,auc_metric = auc(observed,
                                                        prob_pred))
  return(list(error.vec,data.frame(prob_1=prob_pred,prob_0=1-prob_pred,class_pred=class_pred,observed=observed) %>% rename(prob_1=1,prob_0=2)))
} 

# else if(grepl("NN3",model.string)){
#       observed <- as.factor(y)
#       predictorsnn3 <- predictors[,-1]
#       CM <- caret::confusionMatrix(factor(round(predict(model.obj,predictorsnn3)[,1]),levels=c(0,1)),observed)
#     } else if(grepl("NN",model.string)){
#       observed <- as.factor(y)
#       CM <- caret::confusionMatrix(factor(predict(model.obj,newdata=predictors,type="class"),levels=c(0,1)),observed)
#     } else if(model.string=="GLM"){
#       # observed <- as.vector(y)
#       # CM <- caret::confusionMatrix(
#       #   as.factor(predict(model.obj,
#       #                     as.matrix(predictors[, -which(colnames(predictors) ==
#       #                                                     "LPS.Added")]),type="class",s='lambda.min')),
#       #   as.factor(observed))
#       observed <- as.factor(y)
#       CM <- caret::confusionMatrix(as.factor(predict(model.obj,newx = as.matrix(predictors[,-1]),type="class")),
#                             observed)
# 
# roc_auc_fxn <- function(model.string,model.obj,train_data,test.data){
#   predictions<- bind_rows(generate.errors.ML(model.string,model.obj,train_data)[[2]] %>%
#     mutate(datatype = "train")%>% 
#     mutate(auc_metric = as.numeric(auc(observed,prob_1))),
#     generate.errors.ML(model.string,model.obj,test.data)[[2]] %>%
#     mutate(datatype = "test")%>% 
#     mutate(auc_metric = as.numeric(auc(observed,prob_1))))
#   ) %>% 
#     mutate(model= model.string) %>% 
#     select(datatype, auc_metric,model) %>% distinct()
#   return(predictions)
# }


# model.data <- LPS.model.data
# nrep <- 2
train_test_stability.2 <- function(nrep=5,model.data){
  split.vec <- seq(0,1,by=.02)
  split.vec <- split.vec[-c(1:5,(length(split.vec)-4):length(split.vec))]
  # split.vec <- split.vec[c(1,5,10,15)]
all.models <- vector("list",length(split.vec))
names(all.models) <- split.vec
for(i in 1:length(split.vec)){
  prop=split.vec[i]
  model.group <- vector("list", nrep)
  names(model.group) <- 1:nrep
  message(paste("Starting",split.vec[i]))
  for(k in 1:nrep){
    model.obj <- vector("list",4)
    names(model.obj) <- c("split.index",
                          "train.scale",
                          "models",
                          "errors") 
    # Train Test Split
    train_index <- createDataPartition(model.data$LPS.Added,p=prop,list=F)
    model.obj[["split.index"]] <- train_index
    # train_data <- model.data[train_index,]
    # test_data <- model.data[-train_index,]
    
    # Scale Data to training data
    train.min.max <- model.data[train_index,] %>%
      pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
      group_by(Gene) %>%
      summarise(CTmin=min(CT_norm1),
                CTmax=max(CT_norm1)) %>%
      ungroup()
    model.obj[["train.scale"]] <- train.min.max
    
    train_data <- model.data[train_index,] %>%
      pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
      full_join(train.min.max,by="Gene") %>%
      mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
      mutate(CT_scaled=ifelse(is.finite(CT_scaled),CT_scaled,1)) %>% 
      select(LPS.Added,Gene,CT_scaled) %>%
      group_by(Gene) %>%
      mutate(row_num=row_number()) %>%
      pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
      select(-row_num)
    
    test_data <- model.data[-train_index,] %>%
      pivot_longer(!LPS.Added,names_to = "Gene",values_to = "CT_norm1") %>%
      full_join(train.min.max,by="Gene") %>%
      mutate(CT_scaled=(CT_norm1-CTmin)/(CTmax-CTmin)) %>%
      mutate(CT_scaled=ifelse(is.finite(CT_scaled),CT_scaled,1))%>%
      select(LPS.Added,Gene,CT_scaled) %>%
      group_by(Gene) %>%
      mutate(row_num=row_number()) %>%
      pivot_wider(names_from="Gene",values_from="CT_scaled") %>%
      select(-row_num)
    # Train Models 
    # function(model.string,model.obj,predictors)
    mods.list <- vector("list", 6)
    names(mods.list) <- c("rf","nb","svm","glm","nn","nn3")
    
    mod.rf <- randomForest(as.factor(LPS.Added)~., data=train_data,
                           proximity=TRUE,importance=TRUE,ntree=500,keep.forest=TRUE)
    mods.list[["rf"]] <- mod.rf
    
    mod.nb <- naive_bayes(as.logical(LPS.Added) ~ ., data = train_data, usekernel = T, type="raw")
    mods.list[["nb"]] <- mod.nb
    
    mod.svm <-  ksvm(LPS.Added~.,data=train_data,type="C-svc",kernel="vanilladot",prob.model=TRUE,C=.1)
    mods.list[["svm"]] <- mod.svm
    
    # mod.glm <- cv.glmnet(x = as.matrix(train_data[, 
    #                                               -which(colnames(train_data) == 
    #                                                        "LPS.Added")]), 
    #                      y = as.vector(train_data$LPS.Added), 
    #                      type.measure = "class", 
    #                      family = "binomial")
    mod.glm <- glmnet(x=train_data[,-1],
                   y=as.factor(train_data$LPS.Added),
                   family = "binomial",lambda = .01)
    mods.list[["glm"]] <- mod.glm

    mod.nn <- nnet(as.factor(LPS.Added) ~ ., data = train_data ,size=8,linout = FALSE,trace=F,decay=.1)
    mods.list[["nn"]] <- mod.nn
    
    # library(RSNNS)
    mod.nn3 <- mlp(x=train_data[,-1],
                   y=train_data$LPS.Added , size = c(80,20,29))
    mods.list[["nn3"]] <- mod.nn3

    
    model.obj[["models"]] <- mods.list
    
    ## Model Errors
    error_tab <- function(model.string,model.obj,train_data,test_data){
      error_train <- generate.errors.ML(model.string,model.obj,predictors=train_data)[[1]] %>% 
        data.frame(error_val=.) %>% 
        rownames_to_column(var="error_type") %>% 
        mutate(split_data="train")
      error_test <- generate.errors.ML(model.string,model.obj,predictors=test_data)[[1]]%>% 
        data.frame(error_val=.) %>% rownames_to_column(var="error_type") %>% 
        mutate(split_data="test")
      error_mat <- bind_rows(error_train,error_test) %>% mutate(model_type=model.string)
      return(error_mat)
    }
    errors <- bind_rows(error_tab(model.string="RF",model.obj=mod.rf,train_data,test_data),
                        error_tab(model.string="GLM",model.obj=mod.glm,train_data,test_data),
                        error_tab(model.string="SVM",model.obj=mod.svm,train_data,test_data),
                        error_tab(model.string="NN",model.obj=mod.nn,train_data,test_data),
                        error_tab(model.string="NN3",model.obj=mod.nn3,train_data,test_data),
                        error_tab(model.string="NB",model.obj=mod.nb,train_data,test_data)) %>% 
      filter(error_type%in%c("Accuracy","Sensitivity","Specificity","Precision","auc_metric"))
    model.obj[["errors"]] <- errors
    # auc <- bind_rows(error_tab(model.string="RF",model.obj=mod.rf,train_data,test_data),
    #                     error_tab(model.string="GLM",model.obj=mod.glm,train_data,test_data),
    #                     error_tab(model.string="SVM",model.obj=mod.svm,train_data,test_data),
    #                     error_tab(model.string="NN",model.obj=mod.nn,train_data,test_data),
    #                     error_tab(model.string="NN3",model.obj=mod.nn3,train_data,test_data),
    #                     error_tab(model.string="NB",model.obj=mod.nb,train_data,test_data))
    # model.obj[["auc"]]
    # mod.nn
    # mod.glm
    # 
    # Model Errors
    # model.obj[["model"]] <- mod
    model.group[[k]] <- model.obj
    if(k%%25==0){
      message(paste("Progress",split.vec[i],":",k))
    }
  }
  all.models[[i]] <- model.group
  # message(paste("Finished",split.vec[i]))

}
error.data.frame <- data.frame()

for(sp in 1:length(all.models)){
  error.data.frame <- bind_rows(error.data.frame,bind_rows(purrr::map(all.models[[sp]],~.x[["errors"]] %>%
                                             mutate(SplitPerc=as.numeric(names(all.models)[sp]))),
                                .id = "rep"))
}
# save(all.models,error.data.frame,model.data,file = paste0("output/traintest_split_stability_",nrep,".RData"))
return(list(all.models,error.data.frame))
}
suppressMessages({
  run_yn=T
  save_yn=T
for(i in 1:20){
  if(run_yn==T){
  system.time({
  output <- train_test_stability.2(nrep=20,model.data = LPS.model.data)
  })
  error.data.frame <- output[[2]]
  if(save_yn==T){
    write.csv(error.data.frame, 
              paste0(working.dr,"traintest_split_stability_errordf_trans_addnn3_auc_scaled_20_",i,".csv"),row.names = FALSE)
    if(i==1){
      # save(output,file="traintest_split_stability_prot_addnn3_auc_scaled_100_1.RData")
    }
  }
}
}
})


files <- list.files(working.dr,pattern ="trans_addnn3_auc",full.names = T)
files <- files[which(grepl(".csv",files))]
error.data.frame <- data.frame()
for(i in files){
  error.data.frame <- bind_rows(error.data.frame,read.csv(i))
}

ggplot(error.data.frame %>% 
         mutate(error_type=ifelse(error_type=="auc_metric","AUC",error_type)) %>% 
         filter(split_data=="test") %>% 
         group_by(SplitPerc,model_type) %>% 
         mutate(mean_error=mean(error_val,na.rm=T),
                sd_error=sd(error_val,na.rm=T),
                mean_sd_top=min(1,mean_error+sd_error),
                mean_sd_low=mean_error-sd_error) %>% 
         ungroup() %>% 
         filter(SplitPerc>=.1&SplitPerc<=.95),
       # aes(x=SplitPerc,y=1-error_val))+
       aes(x=SplitPerc,y=error_val))+
  geom_point(alpha=.025,size=1)+
  # geom_line(aes(y=1-mean_error),color="forestgreen",size=2)+
  geom_line(aes(y=mean_error),color="forestgreen",size=1.5)+
  # geom_line(aes(y=1-mean_sd_top),color="forestgreen",size=1.5,linetype=3,alpha=.9)+
  # geom_line(aes(y=1-mean_sd_low),color="forestgreen",size=1.5,linetype=3,alpha=.9)+
  # geom_ribbon(aes(ymax = 1-mean_sd_low, ymin = 1-mean_sd_top),linetype=2, alpha = 0.25)+
  # geom_ribbon(aes(ymin = mean_sd_low, ymax = mean_sd_top),linetype=2, alpha = 0.25)+
  xlab("Training Data Percetage")+
  ylab("Testing Error")+
  facet_grid(error_type~model_type)+
  theme_clean()+ 
  scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        )+
  # coord_cartesian(ylim=c(0,.5),xlim=c(.1,.9))
  coord_cartesian(ylim=c(.5,1),xlim=c(.1,.9))

library(ggokabeito)

files <- list.files(working.dr,pattern ="prot_addnn3_auc",full.names = T)
files <- files[which(grepl(".csv",files))]
error.data.frame.prot <- data.frame()
for(i in files){
  error.data.frame.prot <- bind_rows(error.data.frame.prot,read.csv(i))
}
error.data.frame.combo <- bind_rows(error.data.frame %>% 
                                      mutate(DataType = "Transcripts"),
                                    error.data.frame.prot%>% 
                                      mutate(DataType = "Protiens"))

error.data.frame.combo %>% filter(error_type=="Accuracy") %>% 
  filter(split_data=="test") %>% 
  filter(grepl("NN",model_type)) %>% 
  group_by(DataType,SplitPerc,model_type) %>% 
  count(error_val) %>% 
  mutate(Not100Perc=ifelse(error_val<1,TRUE,FALSE)) %>% 
  group_by(DataType,SplitPerc,model_type,Not100Perc,group=Not100Perc) %>% 
  summarise(total_mods=sum(n)) %>% 
  mutate(prop=ifelse(DataType=="Transcripts",total_mods/500,total_mods/500)) %>% 
  filter(!Not100Perc) %>% 
  ggplot(aes(x=SplitPerc,y=prop,color=model_type))+
  geom_line(size=2)+
  facet_wrap(DataType~.)+
  ylab("Proportion of Models")+
  theme_clean()+ 
  scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        )+
  scale_color_okabe_ito()

```


# Hyperparameter Tuning 

##Tune NNET
```{r NN basic tune size}
Acc_breaks = c(seq(.5,0.8,length=100),      #for red
                   seq(0.81,0.95,length=100),  #for white, this puts the key value, 0.05, at the center of the white value range
                   seq(0.96,1,length=100))  

nnetGrid <-  expand.grid(size = seq(from = 1, to = 11, by = 1),#only goes up to 11 before erroring
                         decay = sort(c(2:10 %o% 10^(-1:-4),0.0001),decreasing = TRUE)) 

nnetGrid2 <-  expand.grid(size = seq(from = 1, to = 11, by = 1),
                          decay = c(0,.5,.1,.05,.01,.005,.001,.0005,.0001)) 

#Use predefined reps/folds with unscaled data:####
nn.pre.mod2.foldsreps <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                    method="nnet",tuneGrid = nnetGrid,trace=FALSE,linout=FALSE,trControl = trainControl(index = foldsreps1))
nn.pre.mod2.foldsreps$bestTune
#     size decay
# 206    8  0.02
nn.pre.mod2.foldsreps<-readRDS("NN_unscaled_foldsreps_caretOpt.rds")

nn.pre.mod.unscaled.foldsreps.trim <- nn.pre.mod2.foldsreps$results[which(nn.pre.mod2.foldsreps$results$size <= 11),]#Errors at greater than 11
nn.pre.mod.unscaled.foldsreps.trim <- nn.pre.mod.unscaled.foldsreps.trim[order(nn.pre.mod.unscaled.foldsreps.trim$decay,decreasing = TRUE),]

nn.pre.mod.unscaled.foldsreps.trim$decay.factor <- as.factor(nn.pre.mod.unscaled.foldsreps.trim$decay)
factors.ordered <- levels(nn.pre.mod.unscaled.foldsreps.trim$decay.factor)[levels(nn.pre.mod.unscaled.foldsreps.trim$decay.factor) %>% as.numeric %>% order(decreasing=FALSE)]

nn.pre.mod2.foldsreps.100 <- nn.pre.mod2.foldsreps$results[which.min(1.0 - nn.pre.mod2.foldsreps$results$Accuracy),] #checking that that is the best

#plot with reordered factor levels:
NN_unscaled_foldsreps_hm <-ggplot(data = nn.pre.mod.unscaled.foldsreps.trim,mapping=aes(x=size,y=decay.factor,fill=Accuracy))+geom_tile()+
  scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
  #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
  ylim(factors.ordered)+ylab("NN `decay` values")+xlab("NN # nodes")+scale_x_continuous(breaks=c(1,3,5,7,9,11))+ggtitle("NN with Unscaled Data")+
          theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))

#saveRDS(nn.pre.mod2.foldsreps,"NN_unscaled_foldsreps_caretOpt.rds")
#saveRDS(nn.pre.mod2.foldsreps$finalModel,"NN_unscaled_foldsreps_caretBestModel.rds")

#Use predefined reps/folds with scaled data:####
nn.pre.mod.scaled.foldsreps <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                    method="nnet",tuneGrid = nnetGrid,trace=FALSE,linout=FALSE,trControl = trainControl(index = foldsreps1))

nn.pre.mod.scaled.foldsreps$bestTune
#     size decay
# 230    7 8e-04
# ggplot(data = nn.pre.mod.scaled.foldsreps$results,mapping=aes(x=size,y=log10(decay),fill=Accuracy))+geom_tile()+scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks,to=c(0,1)))

nn.pre.mod.scaled.foldsreps <- readRDS("NN_scaled_foldsreps_caretOpt.rds")

nn.pre.mod.scaled.foldsreps.trim <- nn.pre.mod.scaled.foldsreps$results[which(nn.pre.mod.scaled.foldsreps$results$size <= 11),]#Errors at greater than 11
nn.pre.mod.scaled.foldsreps.trim <- nn.pre.mod.scaled.foldsreps.trim[order(nn.pre.mod.scaled.foldsreps.trim$decay,decreasing = TRUE),]

ggplot(data = nn.pre.mod.scaled.foldsreps.trim,mapping=aes(x=size,y=as.character(decay),fill=Accuracy))+geom_tile()+scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks,to=c(0,1)))

nn.pre.mod.scaled.foldsreps.trim$decay.factor <- as.factor(nn.pre.mod.scaled.foldsreps.trim$decay)
factors.ordered <- levels(nn.pre.mod.scaled.foldsreps.trim$decay.factor)[levels(nn.pre.mod.scaled.foldsreps.trim$decay.factor) %>% as.numeric %>% order(decreasing=FALSE)]

#test on training,testing, and validation data:
NN.unscaled.checktrain.CM <- confusionMatrix(factor(predict(test.mod.nn,newdata=train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],type="class"),levels=c(0,1)),as.factor(train_data_scaled$LPS.Added))

#save things
# saveRDS(nn.pre.mod.scaled.foldsreps,"NN_scaled_foldsreps_caretOpt.rds")
# saveRDS(nn.pre.mod.scaled.foldsreps$finalModel,"NN_scaled_foldsreps_caretBestModel.rds")

#plot with reordered factor levels:
NN_scaled_foldsreps_hm <-ggplot(data = nn.pre.mod.scaled.foldsreps.trim,mapping=aes(x=size,y=decay.factor,fill=Accuracy))+geom_tile()+
  scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
  ylim(factors.ordered)+ylab("NN `decay` values")+xlab("NN # nodes")+scale_x_continuous(breaks=c(1,3,5,7,9,11))+ggtitle("NN with Scaled Data")+
          theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))

```

##Tune Multi-layer Neural net
```{r nn}
set.seed(69)
library(neuralnet)

#caret optimization ####
nnetGrid1 <-  expand.grid(layer1 = seq(from = 20, to = 100, by = 20),
                          layer2 = seq(from = 10, to = 60, by = 10)) 
nnetGrid1p2 <- nnetGrid1[-which(nnetGrid1$layer1 < nnetGrid1$layer2),]
nnetGrid2 <-  expand.grid(layer1 = seq(from = 20, to = 100, by = 20),
                          layer2 = seq(from = 10, to = 60, by = 10),
                          layer3 = seq(from = 5, to = 30, by = 5)) 
nnetGrid2p2 <- nnetGrid2[-which(nnetGrid2$layer1 < nnetGrid2$layer2 | nnetGrid2$layer2 < nnetGrid2$layer3),]
nnetGrid3 <-  expand.grid(layer1 = seq(from = 20, to = 100, by = 10),
                          layer2 = seq(from = 10, to = 60, by = 5),
                          layer3 = seq(from = 5, to = 30, by = 2)) 
##Using 3 layers, "neuralnet" package:####

nn2layer.pre.mod.scaled.foldsreps <- train(LPS.Added~.,data=train_data_scaled,
                    method="neuralnet",threshold = 0.01, err.fct = "sse",tuneGrid = nnetGrid2p2,trControl = trainControl(index = foldsreps1))

nn2layer.pre.mod.scaled.foldsreps$bestTune
#   layer1 layer2 layer3
# 1     20     10      5
##Using 3 layers, "RSNNS" package:####
library(RSNNS)
require(devtools)
library(NeuralNetTools)
library(data.table)
###train at default "learnFuncParams" values####
nn3layer.pre.mod.scaled.foldsreps <- caret::train(x=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled) == "LPS.Added")]),y=as.factor(train_data_scaled$LPS.Added),
                    method="mlpML",tuneGrid = nnetGrid2p2,trControl = trainControl(index = foldsreps1))

nn3layer.pre.mod.scaled.foldsreps$bestTune
#     layer1 layer2 layer3
# 102    100     40     30
saveRDS(nn3layer.pre.mod.scaled.foldsreps$finalModel,"3LayerNN_scaledtranscripts_foldsreps_smallcaretFinalModel.rds")
saveRDS(nn3layer.pre.mod.scaled.foldsreps,"3LayerNN_scaledtranscripts_foldsreps_smallcaret.rds")

###train at default "learnFuncParams" values with finer grid#####
nn3layer2.pre.mod.scaled.foldsreps <- caret::train(x=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled) == "LPS.Added")]),y=as.factor(train_data_scaled$LPS.Added),
                    method="mlpML",tuneGrid = nnetGrid3,trControl = trainControl(index = foldsreps1))

 nn3layer2.pre.mod.scaled.foldsreps$bestTune
#     layer1 layer2 layer3
# 897     80     20     29
 nn3layer2.pre.mod.scaled.foldsreps$results[897,]
#     layer1 layer2 layer3  Accuracy     Kappa AccuracySD    KappaSD
# 897     80     20     29 0.9971429 0.9942857 0.01669328 0.03338657
saveRDS(nn3layer2.pre.mod.scaled.foldsreps$finalModel,"3LayerNN_scaledtranscripts_foldsreps_largercaretFinalModel.rds")
saveRDS(nn3layer2.pre.mod.scaled.foldsreps,"3LayerNN_scaledtranscripts_foldsreps_largercaret.rds")

#Train at learnFuncParams = c(0.01,0) (performed well in some pilot tests)
nn3layer.scaled.foldsreps.LFPp01 <- caret::train(x=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled) == "LPS.Added")]),y=as.factor(train_data_scaled$LPS.Added),
                    method="mlpML",learnFuncParams=c(0.01,0),tuneGrid = nnetGrid2p2,trControl = trainControl(index = foldsreps1))
nn3layer.scaled.foldsreps.LFPp01$bestTune
#     layer1 layer2 layer3
# 113    100     60     25
saveRDS(nn3layer.scaled.foldsreps.LFPp01$finalModel,"3LayerNN_scaledtranscripts_foldsreps_LFP01_smallcaretFinalModel.rds")
saveRDS(nn3layer.scaled.foldsreps.LFPp01,"3LayerNN_scaledtranscripts_foldsreps_LFP01_smallcaret.rds")


#Train at learnFuncParams = c(1,0) (performed best in some pilot tests)
nn3layer.scaled.foldsreps.LFP1 <- caret::train(x=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled) == "LPS.Added")]),y=as.factor(train_data_scaled$LPS.Added),
                    method="mlpML",learnFuncParams=c(1,0),tuneGrid = nnetGrid2p2,trControl = trainControl(index = foldsreps1))
nn3layer.scaled.foldsreps.LFP1$bestTune
#    layer1 layer2 layer3
# 81     80     60     15
saveRDS(nn3layer.scaled.foldsreps.LFP1$finalModel,"3LayerNN_scaledtranscripts_foldsreps_LFP1_smallcaretFinalModel.rds")
saveRDS(nn3layer.scaled.foldsreps.LFP1,"3LayerNN_scaledtranscripts_foldsreps_LFP1_smallcaret.rds")

###exploring importance####
model.nn3opt.Imp <- data.table(olden(nn3layer.pre.mod.scaled.foldsreps$finalModel,bar_plot = FALSE,out_var='Output_1'),"Gene"=rownames(olden(nn3layer.pre.mod.scaled.foldsreps$finalModel,bar_plot = FALSE,out_var='Output_1')))
model.nn3opt.Imp <- model.nn3opt.Imp[order(abs(model.nn3opt.Imp$importance),decreasing = T),]
model.nn3opt.Imp$Gene <- gsub("Input_","",model.nn3opt.Imp$Gene)
write.csv(model.nn3opt.Imp,"3LayerNN_smallopt_scaledTranscript_OldenImp.csv")

###Try changing the "learnFuncParams" value####
library(RSNNS)
model.nn22 <- mlp(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")], train_data_scaled$LPS.Added, size=c(100,40,30), learnFuncParams=c(1,0), 
              maxit=50)

model.nn22.results <- predict(model.nn22,test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")])
compare_nn22 <- data.frame("actual" = test_data_scaled$LPS.Added,"NN_predict3"=model.nn22.results)
nn22_CM3<-caret::confusionMatrix(as.factor(round(compare_nn22[,2],0)), as.factor(test_data_scaled$LPS.Added))
nn22_CM3
#with lower "learFuncParamS"
model.nn202 <- mlp(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")], train_data_scaled$LPS.Added, size=c(100,40,30), learnFuncParams=c(0.01,0), 
              maxit=50)

model.nn202.results <- predict(model.nn202,test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")])
compare_nn202 <- data.frame("actual" = test_data_scaled$LPS.Added,"NN_predict3"=model.nn202.results)
nn202_CM3<-caret::confusionMatrix(as.factor(round(compare_nn202[,2],0)), as.factor(test_data_scaled$LPS.Added))
nn22_CM3

nn3layer.scaled.foldsreps.results <- nn3layer.pre.mod.scaled.foldsreps$results

###plot with reordered factor levels:####
#default "learnFuncParams" values:
nn3layer12_scaled_foldsreps_hm <-ggplot(data = nn3layer.pre.mod.scaled.foldsreps$results,mapping=aes(x=layer1,y=layer2,fill=Accuracy))+geom_tile()+
  scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
  #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
  #ylim(factors.ordered)+
  ylab("Hidden Notes in Layer2")+
  xlab("Hidden Notes in Layer1")+
  #scale_x_continuous(breaks=c(1,3,5,7,9,11))+
  ggtitle("NN with Scaled Transcript Data")+
          theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))
nn3layer12_scaled_foldsreps_hm

#default "learnFuncParams" values with finer grid:
nn3laye2_r12_scaled_foldsreps_hm <-ggplot(data = nn3layer2.pre.mod.scaled.foldsreps$results,mapping=aes(x=layer1,y=layer2,fill=Accuracy))+geom_tile()+
  scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
  #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
  #ylim(factors.ordered)+
  ylab("Hidden Nodes in Layer2")+
  xlab("Hidden Nodes in Layer1")+
  #scale_x_continuous(breaks=c(1,3,5,7,9,11))+
  ggtitle("NN with Scaled Transcript Data")+
          theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))
nn3laye2_r12_scaled_foldsreps_hm

nn3laye2_r13_scaled_foldsreps_hm <-ggplot(data = nn3layer2.pre.mod.scaled.foldsreps$results,mapping=aes(x=layer1,y=layer3,fill=Accuracy))+geom_tile()+
  scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
  #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
  #ylim(factors.ordered)+
  ylab("Hidden Nodes in Layer3")+
  xlab("Hidden Nodes in Layer1")+
  #scale_x_continuous(breaks=c(1,3,5,7,9,11))+
  ggtitle("NN with Scaled Transcript Data")+
          theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))
nn3laye2_r13_scaled_foldsreps_hm

#Changing learnFuncParams to= c(0.01,0) #surprisingly not as good.
nn3layer.scaled.foldsreps.LFPp01_hm <-ggplot(data = nn3layer.scaled.foldsreps.LFPp01$results,mapping=aes(x=layer1,y=layer2,fill=Accuracy))+geom_tile()+
  scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
  #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
  #ylim(factors.ordered)+
  ylab("Hidden Nodes in Layer2")+
  xlab("Hidden Nodes in Layer1")+
  #scale_x_continuous(breaks=c(1,3,5,7,9,11))+
  ggtitle("NN with Scaled Data")+
          theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))

nn3layer.scaled.foldsreps.LFPp01_hm
```



##Tune RF
```{r RF basic tune size}
rfGrid <-  expand.grid(mtry = seq(from = 1, to = 40, by = 1))

#Unscaled Using predefined folds/reps with the data:####
{
  ##25 trees####
  rf.pre.mod25.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 25,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod25.unscaled$bestTune
  #   mtry
  # 17    17
  rf.ntree25.unscaled <- rf.pre.mod25.unscaled$results
  rf.ntree25.unscaled$ntree <- 25
  rf.caretresults.unscaled <- rf.ntree25.unscaled
  ##50 trees####
  rf.pre.mod50.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 50,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod50.unscaled$bestTune
  #   mtry
  # 12    12
  rf.ntree50.unscaled <- rf.pre.mod50.unscaled$results
  rf.ntree50.unscaled$ntree <- 50
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree50.unscaled)
  ##75 trees####
  rf.pre.mod75.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 75,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod75.unscaled$bestTune
  #   mtry
  # 14    14
  rf.ntree75.unscaled <- rf.pre.mod75.unscaled$results
  rf.ntree75.unscaled$ntree <- 75
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree75.unscaled)
  ##100 trees####
  rf.pre.mod100.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 100,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod100.unscaled$bestTune
  #   mtry
  # 6    6
  rf.ntree100.unscaled <- rf.pre.mod100.unscaled$results
  rf.ntree100.unscaled$ntree <- 100
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree100.unscaled)
  ##125 trees####
  rf.pre.mod125.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 125,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod125.unscaled$bestTune
  #   mtry
  # 14    14
  rf.ntree125.unscaled <- rf.pre.mod125.unscaled$results
  rf.ntree125.unscaled$ntree <- 125
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree125.unscaled)
  ##150 trees####
  rf.pre.mod150.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 150,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod150.unscaled$bestTune
  #   mtry
  # 15    15
  rf.ntree150.unscaled <- rf.pre.mod150.unscaled$results
  rf.ntree150.unscaled$ntree <- 150
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree150.unscaled)
  ##175 trees####
  rf.pre.mod175.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 175,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod175.unscaled$bestTune
  #   mtry
  # 7    7
  rf.ntree175.unscaled <- rf.pre.mod175.unscaled$results
  rf.ntree175.unscaled$ntree <- 175
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree175.unscaled)
  ##200 trees####
  rf.pre.mod200.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 200,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod200.unscaled$bestTune
  #   mtry
  # 13    13
  rf.ntree200.unscaled <- rf.pre.mod200.unscaled$results
  rf.ntree200.unscaled$ntree <- 200
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree200.unscaled)
  ##225 trees####
  rf.pre.mod225.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 225,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod225.unscaled$bestTune
  #   mtry
  # 7    7
  rf.ntree225.unscaled <- rf.pre.mod225.unscaled$results
  rf.ntree225.unscaled$ntree <- 225
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree225.unscaled)
  ##250 trees####
  rf.pre.mod250.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 250,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod250.unscaled$bestTune
  #   mtry
  # 18    18
  rf.ntree250.unscaled <- rf.pre.mod250.unscaled$results
  rf.ntree250.unscaled$ntree <- 250
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree250.unscaled)
  ##275 trees####
  rf.pre.mod275.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 275,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod275.unscaled$bestTune
  #   mtry
  # 8    8
  rf.ntree275.unscaled <- rf.pre.mod275.unscaled$results
  rf.ntree275.unscaled$ntree <- 275
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree275.unscaled)
  ##300 trees####
  rf.pre.mod300.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 300,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod300.unscaled$bestTune
  #   mtry
  # 18    18
  rf.ntree300.unscaled <- rf.pre.mod300.unscaled$results
  rf.ntree300.unscaled$ntree <- 300
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree300.unscaled)
  ##325 trees####
  rf.pre.mod325.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 325,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod325.unscaled$bestTune
  #   mtry
  # 12    12
  rf.ntree325.unscaled <- rf.pre.mod325.unscaled$results
  rf.ntree325.unscaled$ntree <- 325
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree325.unscaled)
  ##350 trees####
  rf.pre.mod350.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 350,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod350.unscaled$bestTune
  #   mtry
  # 14    14
  rf.ntree350.unscaled <- rf.pre.mod350.unscaled$results
  rf.ntree350.unscaled$ntree <- 350
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree350.unscaled)
  ##375 trees####
  rf.pre.mod375.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 375,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod375.unscaled$bestTune
  #   mtry
  # 16    16
  rf.ntree375.unscaled <- rf.pre.mod375.unscaled$results
  rf.ntree375.unscaled$ntree <- 375
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree375.unscaled)
  ##400 trees####
  rf.pre.mod400.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 400,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod400.unscaled$bestTune
  #   mtry
  # 17    17
  rf.ntree400.unscaled <- rf.pre.mod400.unscaled$results
  rf.ntree400.unscaled$ntree <- 400
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree400.unscaled)
  ##425 trees####
  rf.pre.mod425.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 425,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod425.unscaled$bestTune
  #   mtry
  # 7    7
  rf.ntree425.unscaled <- rf.pre.mod425.unscaled$results
  rf.ntree425.unscaled$ntree <- 425
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree425.unscaled)
  ##450 trees####
  rf.pre.mod450.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 450,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod450.unscaled$bestTune
  #   mtry
  # 8    8
  rf.ntree450.unscaled <- rf.pre.mod450.unscaled$results
  rf.ntree450.unscaled$ntree <- 450
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree450.unscaled)
  ##475 trees####
  rf.pre.mod475.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 475,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod475.unscaled$bestTune
  #   mtry
  # 5    5
  rf.ntree475.unscaled <- rf.pre.mod475.unscaled$results
  rf.ntree475.unscaled$ntree <- 475
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree475.unscaled)
  ##500 trees####
  rf.pre.mod500.unscaled <- train(x=train_data[,-1],as.factor(train_data$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 500,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod500.unscaled$bestTune
  #   mtry
  # 8    8
  rf.ntree500.unscaled <- rf.pre.mod500.unscaled$results
  rf.ntree500.unscaled$ntree <- 500
  rf.caretresults.unscaled <- rbind (rf.caretresults.unscaled,rf.ntree500.unscaled)
  
}

  #Plotting####
  rf.caretresults.unscaled.foldsreps <- rf.caretresults.unscaled
  rf.caretresults.unscaled.foldsreps$mtry.factor <- as.factor(rf.caretresults.unscaled$mtry)
  rf.factors.ordered <- levels(rf.caretresults.unscaled.foldsreps$mtry.factor)[levels(rf.caretresults.unscaled.foldsreps$mtry.factor) %>% as.numeric %>% order(decreasing=FALSE)]
  #plot with reordered factor levels:
  RF_unscaled_foldsreps_hm <-ggplot(data = rf.caretresults.unscaled.foldsreps,mapping=aes(x=ntree,y=mtry.factor,fill=Accuracy))+geom_tile()+
    scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
    #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
    ylim(rf.factors.ordered)+ylab("RF 'mtry'")+xlab("RF 'ntree'")+
    scale_x_continuous(breaks=c(100,200,300,400,500))+
    ggtitle("RF with Unscaled Data")+
            theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))
  

#Scaled Using predefined folds/reps with the data:####
  {
  ##25 trees####
  rf.pre.mod25.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 25,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod25.scaled$bestTune
  ##Message - unscaled is worse
  #   mtry
  # 19    19
  rf.caretresults <- rf.pre.mod25.scaled$results
  rf.caretresults$ntree <- 25
  ##50 trees####
  rf.pre.mod50.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 50,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod50.scaled$bestTune
  #   mtry
  # 23    23
  rf.ntree50 <- rf.pre.mod50.scaled$results
  rf.ntree50$ntree <- 50
  rf.caretresults <- rbind (rf.caretresults,rf.ntree50)
  ##75 trees####
  rf.pre.mod75.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 75,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod75.scaled$bestTune
  #   mtry
  # 8    8
  rf.ntree75 <- rf.pre.mod75.scaled$results
  rf.ntree75$ntree <- 75
  rf.caretresults <- rbind (rf.caretresults,rf.ntree75)
  ##100 trees####
  rf.pre.mod100.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 100,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod100.scaled$bestTune
  #   mtry
  # 16    16
  rf.ntree100 <- rf.pre.mod100.scaled$results
  rf.ntree100$ntree <- 100
  rf.caretresults <- rbind (rf.caretresults,rf.ntree100)
  ##125 trees####
  rf.pre.mod125.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 125,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod125.scaled$bestTune
  #   mtry
  # 8    8
  rf.ntree125 <- rf.pre.mod125.scaled$results
  rf.ntree125$ntree <- 125
  rf.caretresults <- rbind (rf.caretresults,rf.ntree125)
  ##150 trees####
  rf.pre.mod150.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 150,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod150.scaled$bestTune
  #   mtry
  # 15    15
  rf.ntree150 <- rf.pre.mod150.scaled$results
  rf.ntree150$ntree <- 150
  rf.caretresults <- rbind (rf.caretresults,rf.ntree150)
  ##175 trees####
  rf.pre.mod175.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 175,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod175.scaled$bestTune
  #   mtry
  # 11    11
  rf.ntree175 <- rf.pre.mod175.scaled$results
  rf.ntree175$ntree <- 175
  rf.caretresults <- rbind (rf.caretresults,rf.ntree175)
  ##200 trees####
  rf.pre.mod200.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 200,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod200.scaled$bestTune
  #   mtry
  # 14    14
  rf.ntree200 <- rf.pre.mod200.scaled$results
  rf.ntree200$ntree <- 200
  rf.caretresults <- rbind (rf.caretresults,rf.ntree200)
  ##225 trees####
  rf.pre.mod225.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 225,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod225.scaled$bestTune
  #   mtry
  # 11    11
  rf.ntree225 <- rf.pre.mod225.scaled$results
  rf.ntree225$ntree <- 225
  rf.caretresults <- rbind (rf.caretresults,rf.ntree225)
  ##250 trees####
  rf.pre.mod250.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 250,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod250.scaled$bestTune
  #   mtry
  # 6    6
  rf.ntree250 <- rf.pre.mod250.scaled$results
  rf.ntree250$ntree <- 250
  rf.caretresults <- rbind (rf.caretresults,rf.ntree250)
  ##275 trees####
  rf.pre.mod275.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 275,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod275.scaled$bestTune
  #   mtry
  # 9    9
  rf.ntree275 <- rf.pre.mod275.scaled$results
  rf.ntree275$ntree <- 275
  rf.caretresults <- rbind (rf.caretresults,rf.ntree275)
  ##300 trees####
  rf.pre.mod300.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 300,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod300.scaled$bestTune
  #   mtry
  # 9    9
  rf.ntree300 <- rf.pre.mod300.scaled$results
  rf.ntree300$ntree <- 300
  rf.caretresults <- rbind (rf.caretresults,rf.ntree300)
  ##325 trees####
  rf.pre.mod325.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 325,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod325.scaled$bestTune
  #   mtry
  # 15    15
  rf.ntree325 <- rf.pre.mod325.scaled$results
  rf.ntree325$ntree <- 325
  rf.caretresults <- rbind (rf.caretresults,rf.ntree325)
  ##350 trees####
  rf.pre.mod350.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 350,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod350.scaled$bestTune
  #   mtry
  # 10    10
  rf.ntree350 <- rf.pre.mod350.scaled$results
  rf.ntree350$ntree <- 350
  rf.caretresults <- rbind (rf.caretresults,rf.ntree350)
  ##375 trees####
  rf.pre.mod375.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 375,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod375.scaled$bestTune
  #   mtry
  # 11    11
  rf.ntree375 <- rf.pre.mod375.scaled$results
  rf.ntree375$ntree <- 375
  rf.caretresults <- rbind (rf.caretresults,rf.ntree375)
  ##400 trees####
  rf.pre.mod400.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 400,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod400.scaled$bestTune
  #   mtry
  # 14    14
  rf.ntree400 <- rf.pre.mod400.scaled$results
  rf.ntree400$ntree <- 400
  rf.caretresults <- rbind (rf.caretresults,rf.ntree400)
  ##425 trees####
  rf.pre.mod425.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 425,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod425.scaled$bestTune
  #   mtry
  # 19    19
  rf.ntree425 <- rf.pre.mod425.scaled$results
  rf.ntree425$ntree <- 425
  rf.caretresults <- rbind (rf.caretresults,rf.ntree425)
  ##450 trees####
  rf.pre.mod450.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 450,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod450.scaled$bestTune
  #   mtry
  # 7    7
  rf.ntree450 <- rf.pre.mod450.scaled$results
  rf.ntree450$ntree <- 450
  rf.caretresults <- rbind (rf.caretresults,rf.ntree450)
  ##475 trees####
  rf.pre.mod475.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 475,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod475.scaled$bestTune
  #   mtry
  # 5    5
  
  rf.ntree475 <- rf.pre.mod475.scaled$results
  rf.ntree475$ntree <- 475
  rf.caretresults <- rbind (rf.caretresults,rf.ntree475)
  ##500 trees####
  rf.pre.mod500.scaled <- train(x=train_data_scaled[,-1],as.factor(train_data_scaled$LPS.Added),
                      method="rf",tuneGrid = rfGrid,
                      ntree = 500,
                      trControl = trainControl("repeatedcv", number=5, repeats=50,index = foldsreps1))
  rf.pre.mod500.scaled$bestTune
  #   mtry
  # 4    4
  rf.ntree500 <- rf.pre.mod500.scaled$results
  rf.ntree500$ntree <- 500
  rf.caretresults <- rbind (rf.caretresults,rf.ntree500)
  
  }
  #Plotting things:####
  rf.caretresults.scaled.foldsreps <- rf.caretresults
  rf.caretresults.scaled.foldsreps$mtry.factor <- as.factor(rf.caretresults$mtry)
  rf.factors.ordered <- levels(rf.caretresults.scaled.foldsreps$mtry.factor)[levels(rf.caretresults.scaled.foldsreps$mtry.factor) %>% as.numeric %>% order(decreasing=FALSE)]
    
  
  #plot with reordered factor levels:
  RF_scaled_foldsreps_hm <-ggplot(data = rf.caretresults.scaled.foldsreps,mapping=aes(x=ntree,y=mtry.factor,fill=Accuracy))+geom_tile()+
    scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks2,to=c(0,1)), limits=c(0.5,1))+
    #scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0.9, breaks = Acc_breaks2,limits = c(0.5,1)) +
    ylim(rf.factors.ordered)+ylab("RF 'mtry'")+xlab("RF 'ntree'")+
    #scale_x_continuous(breaks=c(1,3,5,7,9,11))+
    ggtitle("RF with Scaled Data")+
            theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))
  
```
##Tune GLM
```{r GLM basic tune}
# glm.pre.mod <- cv.glmnet(x=as.matrix(LPS.model.data[,
#                                           -which(colnames(LPS.model.data) ==
#                                                      "LPS.Added")]),
#                    y = as.vector(LPS.model.data$LPS.Added),family = "binomial",type.measure = "class",
#                    lambda=c(1, 5e-1,1e-1,5e-2,1e-2,5e-3,1e-3,5e-4,1e-4,5e-5,1e-5))
glm_grid=expand.grid(lambda=sort(c(2:10 %o% 10^(-1:-4),0.0001),decreasing = TRUE),
                     alpha=c(1, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1,0))

glm_grid2=expand.grid(lambda=c(1, 5e-1,1e-1,5e-2,1e-2,5e-3,1e-3,5e-4,1e-4,5e-5,1e-5),
                     alpha=c(1, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1,0))
#Unscaled using definedfoldsreps####
runglm = F
if(runglm == T){
  glm.pre.mod.unscaled.foldsreps <- train(x=train_data[,-c(1)],y=as.factor(train_data$LPS.Added),trControl = trainControl(index = foldsreps1),method="glmnet",family="binomial",tuneGrid=glm_grid)
results.df.trunc <- data.frame(lambda=as.numeric(glm.pre.mod.unscaled.foldsreps$results$lambda),
                               alpha=as.numeric(glm.pre.mod.unscaled.foldsreps$results$alpha),
                               Accuracy=as.numeric(glm.pre.mod.unscaled.foldsreps$results$Accuracy))
}else{
  glm.pre.mod.unscaled.foldsreps <- readRDS("GLMNET_unscaled_foldsreps_caretOpt.rds")
}

glm.pre.mod.unscaled.foldsreps$bestTune
#     alpha lambda
# 399     1    0.2
glm.pre.mod.unscaled.foldsreps.df <- glm.pre.mod.unscaled.foldsreps$results

glm.pre.mod.unscaled.foldsreps.df$lambda.factor <- as.factor(glm.pre.mod.unscaled.foldsreps.df$lambda)
lambda.factors.ordered <- levels(glm.pre.mod.unscaled.foldsreps.df$lambda.factor)[levels(glm.pre.mod.unscaled.foldsreps.df$lambda.factor) %>% as.numeric %>% order(decreasing=FALSE)]

# saveRDS(glm.pre.mod.unscaled.foldsreps,"GLMNET_unscaled_foldsreps_caretOpt.rds")
# saveRDS(glm.pre.mod.unscaled.foldsreps$finalModel,"GLMNET_unscaled_foldsreps_caretBestModel.rds")
glm.pre.mod.unscaled.foldsreps <- readRDS("GLMNET_unscaled_foldsreps_caretOpt.rds")


#Scaled Using Folds/Reps####
runglm_scaled = F
if(runglm_scaled == T){
  glm.pre.mod.scaled.foldsreps <- train(x=train_data_scaled[,-c(1)],y=as.factor(train_data_scaled$LPS.Added),trControl = trainControl(index = foldsreps1),method="glmnet",family="binomial",tuneGrid=glm_grid)
  
  #saveRDS(glm.pre.mod.scaled.foldsreps,"GLMNET_scaled_foldsreps_caretOpt.rds")
#saveRDS(glm.pre.mod.scaled.foldsreps$finalModel,"GLMNET_scaled_foldsreps_caretBestModel.rds")
}else{
  glm.pre.mod.scaled.foldsreps <- readRDS("GLMNET_scaled_foldsreps_caretOpt.rds")
}


glm.pre.mod.scaled.foldsreps$bestTune
#     alpha lambda
# 399     1    0.2
#     alpha lambda
# 318   0.8   0.04
glm.pre.mod.scaled.foldsreps.df <- glm.pre.mod.scaled.foldsreps$results

glm.pre.mod.scaled.foldsreps.df$lambda.factor <- as.factor(glm.pre.mod.scaled.foldsreps.df$lambda)
lambda.factors.ordered <- levels(glm.pre.mod.scaled.foldsreps.df$lambda.factor)[levels(glm.pre.mod.scaled.foldsreps.df$lambda.factor) %>% as.numeric %>% order(decreasing=FALSE)]

###Load Katy's previous folds/reps ####
foldsreps1 <- readRDS("index_save_5_50.rds")
foldsreps1.df <- data.frame(foldsreps1)
foldsrep1.df.1 <- foldsreps1.df[,1:50]
foldsrep1.df.2 <- foldsreps1.df[,51:100]
foldsrep1.df.3 <- foldsreps1.df[,101:150]
foldsrep1.df.4 <- foldsreps1.df[,151:200]
foldsrep1.df.5 <- foldsreps1.df[,201:250]

### Use the caret package to optimize pairs of alpha and lambda for an elastic net:####
model.net <- train(as.factor(LPS.Added) ~., data = train_data_scaled, method = "glmnet",trControl = trainControl("cv", number = 10),family="binomial",tuneLength = 100)

model.net$bestTune #picks: alpha= 0.1, lambda = 0.7067319 (0.5487302 when seed is 400)
coef.opt <- as.data.frame(as.matrix(coef(model.net$finalModel, s=model.net$bestTune$lambda))[-1,])
colnames(coef.opt) <- "coefs"
coef.opt$vars <- rownames(coef.opt)
coef.opt.nz <-  coef.opt[which(coef.opt$coefs !=0),]
coef.opt.nz.ordered <- as.data.frame(coef.opt.nz[order(abs(coef.opt.nz$coefs),decreasing = TRUE),])

confusionMatrix(as.factor(round(predict(model.net$finalModel,as.matrix(test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")]),s=model.net$bestTune$lambda,type="response"),0)),as.factor(test_data_scaled$LPS.Added))

saveRDS(model.net,"GLMNET_caretopt.rds")

### Graph Results from caret optimization, pick/test alternative parameter values####
#Check selected optimal pair, but from plots, looks like would want to pick alpha < 0.4, lambda < 0.3 (< 0.2 for seed of 400)
scatterplot3d(model.net$results$alpha,model.net$results$lambda,model.net$results$Accuracy, angle = 55)
plot(model.net$results$alpha,model.net$results$Accuracy)
plot(model.net$results$lambda,model.net$results$Accuracy)

#From plot, pick an alpha/labmda pair and test:
lambda.select <- model.net$results$lambda[which.min(abs(0.2-model.net$results$lambda[which(model.net$results$alpha == 0.4)]))]

test3 <- glmnet(as.matrix(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")]),as.vector(train_data_scaled$LPS.Added),family="binomial",alpha = 0.4)
coef.test3 <- as.data.frame(as.matrix(coef(test3, s=lambda.select))[-1,]) #works very well when the seed is set at 400
colnames(coef.test3) <- "coefs"
coef.test3$vars <- rownames(coef.test3)
coef.test3.nz <-  coef.test3[which(coef.test3$coefs !=0),]
coef.test3.nz.ordered <- as.data.frame(coef.test3.nz[order(abs(coef.test3.nz$coefs),decreasing = TRUE),])

confusionMatrix(as.factor(round(predict(test3,as.matrix(test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")]),s=lambda.select,type="response"),0)),as.factor(test_data_scaled$LPS.Added))

#Seed = 1306, alpha = 0.4, lambda = 0.2971783
#             coefs  vars
# CCL5  -2.72132867  CCL5
# CXCL2 -1.46181320 CXCL2
# LTB   -1.29510708   LTB
# CXCL1 -1.24846110 CXCL1
# IL6   -0.55479476   IL6
# CCL2  -0.52960891  CCL2
# CSF3  -0.47984552  CSF3
# CCL20 -0.47331357 CCL20
# CXCL5 -0.34464967 CXCL5
# VEGFA  0.09518796 VEGFA
# IL18   0.09423151  IL18
# BMP4   0.07277107  BMP4

#Seed = 1306, alpha = 0.4, lambda = 0.1971511
#              coefs  vars
# CCL5  -2.750503622  CCL5
# CXCL2 -1.470014044 CXCL2
# LTB   -1.306445001   LTB
# CXCL1 -1.257973509 CXCL1
# IL6   -0.562318756   IL6
# CCL2  -0.535858976  CCL2
# CSF3  -0.483224198  CSF3
# CCL20 -0.477447402 CCL20
# CXCL5 -0.346694101 CXCL5
# VEGFA  0.106058884 VEGFA
# IL18   0.102073710  IL18
# BMP4   0.084644078  BMP4
# IL22   0.005629512  IL22
# IL1B  -0.004392623  IL1B

#Seed = 400
#           coefs  vars
# CCL5  -2.7268124  CCL5
# CXCL2 -1.5044520 CXCL2
# CXCL1 -1.4203935 CXCL1
# IL6   -0.6712120   IL6
# CCL20 -0.6167762 CCL20
# LTB   -0.6116767   LTB
# CCL2  -0.4302746  CCL2
# CSF3  -0.2097995  CSF3
# CXCL8 -0.2048959 CXCL8


### Custom glmnet function to use given folds, optimize lambda values over repeats ####
glmnet.lambdaopt <- function(x,y, testx=NULL,testy=NULL,repeats = 50, folds = 5,foldsreps = foldsreps1.df,alpha.set=0.4,lambdas = c(seq(0.000001,0.000009,0.000001),seq(0.00001,0.00009,0.00001),seq(0.0001,0.0009,0.0001),seq(0.001,0.009,0.0002),seq(0.01,0.09,0.002),seq(0.1,0.9,0.02))){
  lambdas = lambdas[order(lambdas,decreasing = TRUE)]
  modelinfo <- NULL
  varimp <- vector(mode='list', length=repeats)
  varimp.scaled <- vector(mode='list', length=repeats)
  models <- vector(mode='list', length=repeats)
  #colnames(results) <- c("Variables","VariablesSD","Accuracy","AccuracySD","Kappa","KappaSD","NonZeroCoeffs","NonZeroCoeffsSD")
  
    #structuring input data (because some of the algorithms take x and y as part of a dataframe, others as two separate arguments)
  data <- x
  data$y <- y
  
   #If testing data is given, structuring test data (because some of the algorithms take x and y as part of a dataframe, others as two separate arguments)
  if(!is.null(testx) || !is.null(testy)){
    testing <- testx
    testing$y <- testy
  }
  
  FoldAccuracies <- data.frame(matrix(ncol=length(sizes),nrow=repeats*folds))
  colnames(FoldAccuracies) <- sizes
  allcoefs<- data.frame("var"=colnames(x))
  allcoefs.scaled<- data.frame("var"=colnames(x))
  for(k in 1:repeats){
    varimp[[k]] <- vector(mode='list', length=folds)
    varimp.scaled[[k]] <- vector(mode='list', length=folds)
    models[[k]] <- vector(mode='list', length=folds)
    kcoefs <- data.frame("var"=colnames(x))
    kcoefs.scaled <- data.frame("var"=colnames(x))
    for(j in 1:folds){
        varimp[[k]][[j]] <- vector(mode='list', length=length(lambdas))
        varimp.scaled[[k]][[j]] <- vector(mode='list', length=length(lambdas))
        models[[k]][[j]] <- vector(mode='list', length=length(lambdas))
        jfoldindex <- foldsreps[,(k-1)*5 + j] #Folds/Reps are ordered by Reps and then by folds.
        jtrain_data <- data[jfoldindex,]
        jtest_data <- data[-jfoldindex,]
        
      newentry <- data.frame(matrix(ncol=4,nrow=length(lambdas)))
      colnames(newentry) <- c("Lambda","Accuracy","Kappa","NonZeroCoeffs")
      jcoefs <- data.frame("var"=colnames(x))
      jcoefs.scaled <- data.frame("var"=colnames(x))
    
      
      imodel <- glmnet(x = as.matrix(jtrain_data[,-which(colnames(jtrain_data)=="y")]),y=as.vector(jtrain_data$y),family = "binomial",lambda = lambdas,alpha = alpha.set) #The glmnet fits a model at each value of lambda. We will then pull out this model at each value.
      models[[k]][[j]]<- imodel
      
      for(i in 1:length(lambdas)){
        check <- confusionMatrix(as.factor(round(predict(imodel,newx=as.matrix(jtest_data[,-which(colnames(jtest_data)=="y")]),type="response",s=lambdas[i]),0)),as.factor(jtest_data$y))
      
       newentry$Lambda[i] <- lambdas[i]
       newentry$Accuracy[i] <- check$overall["Accuracy"]
       newentry$Kappa[i] <- check$overall["Kappa"]
      
      coefs <- as.data.frame(as.matrix((coef(imodel,s=lambdas[i]))))
      coefs2 <- coefs[rownames(coefs) != "(Intercept)",]
      names(coefs2) <- rownames(coefs)[rownames(coefs)!= "(Intercept)"]
      coefs2[is.na(coefs2)] <- 0
      vimp <- data.frame( unname(coefs2), var = names(coefs2))
      colnames(vimp)[1]<-paste0("Lambda",lambdas[i],".",k,".",j)
      rownames(vimp) <- names(coefs2)
      vimp <- vimp[order(abs(vimp[,1]), decreasing = TRUE), , drop = FALSE]
      vimpscaled <- vimp
      vimpscaled[,1] <- as.numeric(abs(vimpscaled[,1])/max(abs(vimpscaled[,1])))
      iNZs <- which(vimpscaled[,1] != 0)
      
      newentry$NonZeroCoeffs[i] <- length(iNZs)
      
      #Save variable importance and model structure for each model
      varimp[[k]][[j]][[i]] <- vimp
      varimp.scaled[[k]][[j]][[i]] <- vimpscaled

      # #Save each model accuracy and kappa
      # models[[k]][[j]][[i]]$AccFold <- check$overall["Accuracy"]
      # models[[k]][[j]][[i]]$Kappa <- check$overall["Kappa"]
      # models[[k]][[j]][[i]]$NonZeroCoeff <- length(iNZs)
      
      joined <- left_join(jcoefs,vimp,by="var")
      jcoefs <- joined
      
      joined.scaled <- left_join(jcoefs.scaled,vimpscaled,by="var")
      jcoefs.scaled <- vimpscaled
      }
      newkcoefs <- left_join(kcoefs,jcoefs,by="var")
      kcoefs <- newkcoefs
      
      newkcoefs.scaled <- left_join(kcoefs.scaled,jcoefs.scaled,by="var")
      kcoefs.scaled <- newkcoefs.scaled
      
      modelinfo <- rbind(modelinfo,newentry)
    }
    newallcoefs <- left_join(allcoefs,kcoefs,by="var")
    allcoefs <- newallcoefs
    
    newallcoefs.scaled <- left_join(allcoefs.scaled,kcoefs.scaled,by="var")
    allcoefs.scaled <- newallcoefs.scaled
  }
  lambdamods.df <- data.frame(matrix(nrow=length(lambdas),ncol=5))
  colnames(lambdamods.df) <- c("Lambda","AccAve","AccSD","NVarAve","NVarSD")
  for(i in 1:length(lambdas)){
    ilambdainfo <- modelinfo[which(modelinfo$Lambda == lambdas[i]),]
    AvgInfo <- MeanSD(ilambdainfo)
    aveinfo2 <- data.frame("Ave" = AvgInfo[1:(length(AvgInfo)/2)],"SD" =AvgInfo[(length(AvgInfo)/2 + 1):length(AvgInfo)])
    lambdamods.df$Lambda[i] <- lambdas[i]
    lambdamods.df$AccAve[i] <- aveinfo2$Ave[which(rownames(aveinfo2)=="Accuracy")]
    lambdamods.df$AccSD[i] <- aveinfo2$SD[which(rownames(aveinfo2)=="Accuracy")]
    lambdamods.df$NVarAve[i] <- aveinfo2$Ave[which(rownames(aveinfo2)=="NonZeroCoeffs")]
    lambdamods.df$NVarSD[i] <- aveinfo2$SD[which(rownames(aveinfo2)=="NonZeroCoeffs")]
  }
  #Use coeff values to determine importance order in unscaled coefficients
  allcoefs2 <- allcoefs
  rownames(allcoefs2) <- allcoefs$var
  allcoefs2 <- allcoefs2[,-which(colnames(allcoefs2)=="var")]
  AvgCoeffs <- data.frame("Ave" = rowSums(allcoefs2)/ncol(allcoefs2), "SD" = apply(allcoefs2,1,sd))
  AvgCoeffs <- AvgCoeffs[order(abs(AvgCoeffs$Ave),decreasing = TRUE),]
  
  #Use coeff values to determine importance order in SCALED coefficients
  allcoefs2.scaled <- allcoefs.scaled
  rownames(allcoefs2.scaled) <- allcoefs.scaled$var
  allcoefs2.scaled <- allcoefs2.scaled[,-which(colnames(allcoefs2.scaled)=="var")]
  AvgCoeffs.scaled <- data.frame("Ave" = rowSums(allcoefs2.scaled)/ncol(allcoefs2.scaled), "SD" = apply(allcoefs2.scaled,1,sd))
  AvgCoeffs.scaled <- AvgCoeffs.scaled[order(abs(AvgCoeffs.scaled$Ave),decreasing = TRUE),]
  
#Pick the optimal value of labmda
lambda.opt <- lambdamods.df$Lambda[which(lambdamods.df$AccAve == max(lambdamods.df$AccAve))]
if(length(lambda.opt) > 1){ #if multiple values of lambda have the same accuracy, pick the one that uses the fewest variables
  lambda.opt.NVar <- lambdamods.df$NVarAve[which(lambdamods.df$Lambda %in% lambda.opt)]
  lambda.opt <- lambdamods.df$Lambda[which(lambdamods.df$AccAve == max(lambdamods.df$AccAve) & lambdamods.df$NVarAve == min(lambda.opt.NVar) )]
}

  return(list("LambdasAll"=lambdamods.df,"lambda.min"=lambda.opt,"modelsinfo"=modelinfo,"avgCoeff"=AvgCoeffs, "AllCoeff"=allcoefs,"avgScaledCoeff"=AvgCoeffs.scaled,"AllScaledCoeff"=allcoefs.scaled,"GLMNetImp"=varimp,"GLMNetImp.scaled"=varimp.scaled,"GLMNetmodels"=models))
}

##Use manually picked alpha of 0.4:
glmnet.alpha0p4.lambdaopt <- glmnet.lambdaopt(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsreps1.df,repeats=50,folds=5,alpha.set=0.4)

glmnet.alpha0p4.lambdamin <- glmnet.alpha0p4.lambdaopt$lambda.min
glmnet.alpha0p4.AvgScaledCoef <- glmnet.alpha0p4.lambdaopt$avgScaledCoeff
glmnet.alpha0p4.OptSummary <- glmnet.alpha0p4.lambdaopt$LambdasAll

ggplot(glmnet.alpha0p4.OptSummary,aes(`Lambda`,`AccAve`))+geom_point()+theme_bw()+ylim(0.99,1.0)
ggplot(glmnet.alpha0p4.OptSummary,aes(`NVarAve`,`AccAve`),ylim(0.99,1.0))+geom_point()+theme_bw()

saveRDS(glmnet.alpha0p4.lambdaopt,"GLMNET_alpha0p4.rds")

##Use caret-selected alpha of 0.1:
glmnet.alpha0p1.lambdaopt <- glmnet.lambdaopt(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsreps1.df,repeats=50,folds=5,alpha.set=0.1)

glmnet.alpha0p1.lambdamin <- glmnet.alpha0p1.lambdaopt$lambda.min
glmnet.alpha0p1.AvgScaledCoef <- glmnet.alpha0p1.lambdaopt$avgScaledCoeff
glmnet.alpha0p1.OptSummary <- glmnet.alpha0p1.lambdaopt$LambdasAll

plot(glmnet.alpha0p1.OptSummary$Lambda,glmnet.alpha0p1.OptSummary$AccAve)
plot(glmnet.alpha0p1.OptSummary$NVarAve,glmnet.alpha0p1.OptSummary$AccAve)

saveRDS(glmnet.alpha0p1.lambdaopt,"GLMNET_alpha0p1.rds")
write_csv(glmnet.alpha0p1.OptSummary,"GLMNET_alpha0p1_OptimizationSummary.csv")

##Test shooting the gap at alpha = 0.25:
glmnet.alpha0p25.lambdaopt <- glmnet.lambdaopt(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsreps1.df,repeats=50,folds=5,alpha.set=0.25)

glmnet.alpha0p25.lambdamin <- glmnet.alpha0p25.lambdaopt$lambda.min
glmnet.alpha0p25.AvgScaledCoef <- glmnet.alpha0p25.lambdaopt$avgScaledCoeff
glmnet.alpha0p25.OptSummary <- glmnet.alpha0p25.lambdaopt$LambdasAll

plot(glmnet.alpha0p25.OptSummary$Lambda,glmnet.alpha0p25.OptSummary$AccAve)
plot(glmnet.alpha0p25.OptSummary$NVarAve,glmnet.alpha0p25.OptSummary$AccAve)

saveRDS(glmnet.alpha0p1.lambdaopt,"GLMNET_alpha0p1.rds")
glmnet.alpha0p1.lambdaopt <- readRDS("glmnet.alpha0p1.lambdaopt")

#Plot together for comparison
ggplot(glmnet.alpha0p4.OptSummary,aes(`Lambda`,`AccAve`))+geom_point()+theme_bw()+geom_point(aes(glmnet.alpha0p1.OptSummary$Lambda,glmnet.alpha0p1.OptSummary$AccAve),color = "blue")+geom_point(aes(glmnet.alpha0p25.OptSummary$Lambda,glmnet.alpha0p25.OptSummary$AccAve),color = "green")+ylim(0.99,1.0)

ggplot(glmnet.alpha0p4.OptSummary,aes(`NVarAve`,`AccAve`))+geom_point()+theme_bw()+geom_point(aes(glmnet.alpha0p1.OptSummary$NVarAve,glmnet.alpha0p1.OptSummary$AccAve),color = "blue")+geom_point(aes(glmnet.alpha0p25.OptSummary$NVarAve,glmnet.alpha0p25.OptSummary$AccAve),color = "green")+ylim(0.99,1.0)


### Average coefficients over the folds/reptitions models with selected alpha and lambda values: ####
allcoef.alpha0p1 <- glmnet.alpha0p1.lambdaopt$AllCoeff
coeff.alpha0p1.lambda0p44 <- data.frame(cbind("var"=allcoef.alpha0p1$var,allcoef.alpha0p1[,grep("Lambda0.44",colnames(allcoef.alpha0p1))]))
rownames(coeff.alpha0p1.lambda0p44)<-coeff.alpha0p1.lambda0p44$var

## Average raw coefficient value
AvgCoeff_a0.1_l0.44 <- data.frame(cbind(coeff.alpha0p1.lambda0p44$var,
                             "AvgCoeff" = rowSums(coeff.alpha0p1.lambda0p44[,-which(colnames(coeff.alpha0p1.lambda0p44) == "var")])/(ncol(coeff.alpha0p1.lambda0p44)-1),
                             "CoeffSD" = apply(coeff.alpha0p1.lambda0p44[,-which(colnames(coeff.alpha0p1.lambda0p44) == "var")],1,sd)))
AvgCoeff_a0.1_l0.44$AvgCoeff <- as.numeric(AvgCoeff_a0.1_l0.44$AvgCoeff)

AvgCoeff_a0.1_l0.44$CoeffSD <- as.numeric(AvgCoeff_a0.1_l0.44$CoeffSD)
AvgCoeff_a0.1_l0.44 <- AvgCoeff_a0.1_l0.44[order(abs(AvgCoeff_a0.1_l0.44$AvgCoeff),decreasing = TRUE),]

write.csv(AvgCoeff_a0.1_l0.44,"GLMNet_AvgCoeffs_ap1_lp44.csv")
write.csv(coeff.alpha0p1.lambda0p44,"GLMNet_Coeffs_ap1_lp44.csv")

## Average scaled coefficient value
norm_to_max <-function(x){
  normx <- x/max(abs(x))
  return(normx)
}
coeff.alpha0p1.lambda0p44.scaled <- as.data.frame(cbind("var"=coeff.alpha0p1.lambda0p44$var,apply(coeff.alpha0p1.lambda0p44[,-which(colnames(coeff.alpha0p1.lambda0p44)=="var")],2,norm_to_max)))
coeff.alpha0p1.lambda0p44.scaled[,-which(colnames(coeff.alpha0p1.lambda0p44)=="var")] <- apply(coeff.alpha0p1.lambda0p44.scaled[,-which(colnames(coeff.alpha0p1.lambda0p44.scaled)=="var")],2,as.numeric)

AvgCoeff_a0.1_l0.44.scaled <- data.frame(cbind(coeff.alpha0p1.lambda0p44.scaled$var,
                             "AvgCoeff" = rowSums(coeff.alpha0p1.lambda0p44.scaled[,-which(colnames(coeff.alpha0p1.lambda0p44.scaled) == "var")])/(ncol(coeff.alpha0p1.lambda0p44.scaled)-1),
                             "CoeffSD" = apply(coeff.alpha0p1.lambda0p44.scaled[,-which(colnames(coeff.alpha0p1.lambda0p44.scaled) == "var")],1,sd)))

AvgCoeff_a0.1_l0.44.scaled$AvgCoeff <- as.numeric(AvgCoeff_a0.1_l0.44.scaled$AvgCoeff)
AvgCoeff_a0.1_l0.44.scaled$CoeffSD <- as.numeric(AvgCoeff_a0.1_l0.44.scaled$CoeffSD)

AvgCoeff_a0.1_l0.44.scaled <- AvgCoeff_a0.1_l0.44.scaled[order(abs(AvgCoeff_a0.1_l0.44.scaled$AvgCoeff),decreasing = TRUE),]

write.csv(AvgCoeff_a0.1_l0.44.scaled,"GLMNet_AvgScaledCoeffs_ap1_lp44.csv")
write.csv(coeff.alpha0p1.lambda0p44.scaled,"GLMNet_ScaledCoeffs_ap1_lp44.csv")

## Average rank of each variable
coeff.alpha0p1.lambda0p44.rank <- coeff.alpha0p1.lambda0p44.scaled
coeff.alpha0p1.lambda0p44.rank[coeff.alpha0p1.lambda0p44.rank == 0] <- NA #make 0s NA so they don't mess with the ranking
coeff.alpha0p1.lambda0p44.rank[,-which(colnames(coeff.alpha0p1.lambda0p44.rank)=="var")] <- coeff.alpha0p1.lambda0p44.rank[,-which(colnames(coeff.alpha0p1.lambda0p44.rank)=="var")] %>% apply(.,2,rank,na.last="keep")
coeff.alpha0p1.lambda0p44.rank[is.na(coeff.alpha0p1.lambda0p44.rank)] <- 85

AvgCoeff_a0.1_l0.44.rank <- data.frame(cbind(coeff.alpha0p1.lambda0p44.rank$var,
                             "AvgRank" = rowSums(coeff.alpha0p1.lambda0p44.rank[,-which(colnames(coeff.alpha0p1.lambda0p44.rank) == "var")])/(ncol(coeff.alpha0p1.lambda0p44.rank)-1),
                             "RankSD" = apply(coeff.alpha0p1.lambda0p44.rank[,-which(colnames(coeff.alpha0p1.lambda0p44.rank) == "var")],1,sd)))

AvgCoeff_a0.1_l0.44.rank$AvgRank <- as.numeric(AvgCoeff_a0.1_l0.44.rank$AvgRank)
AvgCoeff_a0.1_l0.44.rank$RankSD <- as.numeric(AvgCoeff_a0.1_l0.44.rank$RankSD)

AvgCoeff_a0.1_l0.44.rank <- AvgCoeff_a0.1_l0.44.rank[order(abs(AvgCoeff_a0.1_l0.44.rank$AvgRank)),]

write.csv(AvgCoeff_a0.1_l0.44.rank,"GLMNet_AvgCoeffsRank_ap1_lp44.csv")
write.csv(coeff.alpha0p1.lambda0p44.rank,"GLMNet_CoeffsRanks_ap1_lp44.csv")

#Number of non-zero coefficients at each fold:
nz.coeff.alpha0p1.lambda0p44 <- data.frame(matrix(nrow = ncol(coeff.alpha0p1.lambda0p44)-1,ncol=4))
colnames(nz.coeff.alpha0p1.lambda0p44) <- c("ModelIndex","Rep","Fold","NonZeroCoeffs")
nz.coeff.alpha0p1.lambda0p44$ModelIndex <- seq(1,250,1)
Reps.Folds <- colnames(coeff.alpha0p1.lambda0p44) %>% gsub("Lambda0.44.","",.)
for(i in 2:ncol(coeff.alpha0p1.lambda0p44)){
  nz.coeff.alpha0p1.lambda0p44$Fold[i-1] <- str_sub(Reps.Folds[i],str_length(Reps.Folds[i]))
  nz.coeff.alpha0p1.lambda0p44$Rep[i-1] <- str_sub(Reps.Folds[i],end=str_length(Reps.Folds[i])-2)
  nz.coeff.alpha0p1.lambda0p44$NonZeroCoeffs[i-1] <- length(which(coeff.alpha0p1.lambda0p44[,i] != 0))
  }
write.csv(nz.coeff.alpha0p1.lambda0p44,"GLMNet_NonZeroCoeff_ap1_lp44.csv")

#### Generate a final model from the entire training set using the lambda value picked earlier: ####
###alpha = 0.4
glmnet.finalmodel.ap4.lp18 <- glmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,family="binomial",alpha=0.4)
glmnet.finalmodel.ap4.lp18$lambdaopt <- glmnet.alpha0p4.lambdaopt$lambda.min

  checktraining <- predict(glmnet.finalmodel.ap4.lp18,newx=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap4.lp18$lambdaopt,type="class")
  trainingCM.ap4.lp18 <- confusionMatrix(as.factor((checktraining)),as.factor(train_data_scaled$LPS.Added))
  
    checktesting <- predict(glmnet.finalmodel.ap4.lp18,newx=as.matrix(test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap4.lp18$lambdaopt,type="class")
  testingCM.ap4.lp18 <- confusionMatrix(as.factor((checktesting)),as.factor(test_data_scaled$LPS.Added))
  
  checkMJLPS <- predict(glmnet.finalmodel.ap4.lp18,newx=as.matrix(MJ.LPS_data_scaled[,-which(colnames(MJ.LPS_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap4.lp18$lambdaopt,type="class")
  MJLPS.CM.ap4.lp18 <- confusionMatrix(as.factor((checkMJLPS)),as.factor(MJ.LPS_data_scaled$LPS.Added))
  
  coefs.glmnet.finalmodel.ap4.lp18 <- as.data.frame(as.matrix(coef(glmnet.finalmodel.ap4.lp18,s=glmnet.finalmodel.ap4.lp18$lambdaopt)))
  coefs.glmnet.finalmodel.ap4.lp18$var <- rownames(coefs.glmnet.finalmodel.ap4.lp18)
  coefsorder.glmnet.finalmodel.ap4.lp18 <- coefs.glmnet.finalmodel.ap4.lp18[order(abs(coefs.glmnet.finalmodel.ap4.lp18$s1),decreasing = TRUE),]
  
  
saveRDS(glmnet.finalmodel.ap4.lp18,"GLMNET_FinalModel_ap4_lp18.rds")

###alpha=0.1
glmnet.finalmodel.ap1.lp44 <- glmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,family="binomial",alpha=0.1)
glmnet.finalmodel.ap1.lp44$lambdaopt <- glmnet.alpha0p1.lambdaopt$lambda.min

  checktraining <- predict(glmnet.finalmodel.ap1.lp44,newx=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap1.lp44$lambdaopt,type="class")
  trainingCM.ap1.lp44 <- confusionMatrix(as.factor(checktraining),as.factor(train_data_scaled$LPS.Added))
  
    checktraining2 <- predict(glmnet.finalmodel.ap1.lp44,newx=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap1.lp44$lambdaopt,type="class")
  trainingCM.ap1.lp44 <- confusionMatrix(as.factor((checktraining2)),as.factor(train_data_scaled$LPS.Added))
  
    checktesting <- predict(glmnet.finalmodel.ap1.lp44,newx=as.matrix(test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap1.lp44$lambdaopt,type="class")
  testingCM.ap1.lp44 <- confusionMatrix(as.factor((checktesting)),as.factor(test_data_scaled$LPS.Added))
  
    checkMJLPS <- predict(glmnet.finalmodel.ap1.lp44,newx=as.matrix(MJ.LPS_data_scaled[,-which(colnames(MJ.LPS_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap1.lp44$lambdaopt,type="class")
  MJLPS.CM.ap1.lp44 <- confusionMatrix(as.factor((checkMJLPS)),as.factor(MJ.LPS_data_scaled$LPS.Added))
  
  #Use coefficient values to show importance:
  coefs.glmnet.finalmodel.ap1.lp44 <- as.data.frame(as.matrix(coef(glmnet.finalmodel.ap1.lp44,s=glmnet.finalmodel.ap1.lp44$lambdaopt)))
  coefs.glmnet.finalmodel.ap1.lp44$var <- rownames(coefs.glmnet.finalmodel.ap1.lp44)
  coefsorder.glmnet.finalmodel.ap1.lp44 <- coefs.glmnet.finalmodel.ap1.lp44[order(abs(coefs.glmnet.finalmodel.ap1.lp44$s1),decreasing = TRUE),]
  
  #Check against the caret package "varImp:"
  glmnet.ap1.lp44.caretImp <- varImp(glmnet.finalmodel.ap1.lp44,lambda = 0.44)
  glmnet.ap1.lp44.caretImp$var <- rownames(glmnet.ap1.lp44.caretImp)
  glmnet.ap1.lp44.caretImp <- glmnet.ap1.lp44.caretImp[order(glmnet.ap1.lp44.caretImp$Overall,decreasing = TRUE),]
  
  #saving things
  saveRDS(glmnet.finalmodel.ap1.lp44,"GLMNET_FinalModel_ap1_lp44.rds")
  write_csv(coefsorder.glmnet.finalmodel.ap1.lp44,"GLMNET_FinalModel_CoefOrder.csv")
  write_csv(glmnet.ap1.lp44.caretImp,"GLMNET_FinalModel_CaretImp.csv")
  
  ###alpha=0.25
glmnet.finalmodel.ap25.lp28 <- glmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,family="binomial",alpha=0.25)
glmnet.finalmodel.ap25.lp28$lambdaopt <- glmnet.alpha0p25.lambdaopt$lambda.min

  checktraining <- predict(glmnet.finalmodel.ap25.lp28,newx=as.matrix(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap25.lp28$lambdaopt,type="response")
  trainingCM.ap25.lp28 <- confusionMatrix(as.factor(round(checktraining,0)),as.factor(train_data_scaled$LPS.Added))
  
    checktesting <- predict(glmnet.finalmodel.ap25.lp28,newx=as.matrix(test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap25.lp28$lambdaopt,type="response")
  testingCM.ap25.lp28 <- confusionMatrix(as.factor(round(checktesting,0)),as.factor(test_data_scaled$LPS.Added))
  
    checkMJLPS <- predict(glmnet.finalmodel.ap25.lp28,newx=as.matrix(MJ.LPS_data_scaled[,-which(colnames(MJ.LPS_data_scaled)=="LPS.Added")]),s = glmnet.finalmodel.ap25.lp28$lambdaopt,type="class")
  MJLPS.CM.ap25.lp28 <- confusionMatrix(as.factor((checkMJLPS)),as.factor(MJ.LPS_data_scaled$LPS.Added))
  
  coefs.glmnet.finalmodel.ap25.lp28 <- as.data.frame(as.matrix(coef(glmnet.finalmodel.ap25.lp28,s=glmnet.finalmodel.ap25.lp28$lambdaopt)))
  coefs.glmnet.finalmodel.ap25.lp28$var <- rownames(coefs.glmnet.finalmodel.ap25.lp28)
  coefsorder.glmnet.finalmodel.ap25.lp28 <- coefs.glmnet.finalmodel.ap25.lp28[order(abs(coefs.glmnet.finalmodel.ap25.lp28$s1),decreasing = TRUE),]
  
  saveRDS(glmnet.finalmodel.ap25.lp28,"GLMNET_FinalModel_ap25_lp28.rds")

# glmnetimportance <- as.data.frame(rfeglmnet_results1[["avgScaledCoeff"]])
# glmnetAveAcc <- as.data.frame(rfeglmnet_results1[["AveAcc"]])
# glmnetAllAcc <-  as.data.frame(rfeglmnet_results1[["AllAcc"]])
# glmnetAllScaledCoeff <- rfeglmnet_results1[["AllScaledCoeff"]]

# #split up into sublists
# rfeglmnet_results1.1 <- rfeglmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsrep1.df.1,repeats=10,folds=5)
# rfeglmnet_results1.1[["avgScaledCoeff"]] 
# 
# rfeglmnet_results1.2 <- rfeglmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsrep1.df.2,repeats=10,folds=5)
# rfeglmnet_results1.2[["avgScaledCoeff"]] 
# 
# rfeglmnet_results1.3 <- rfeglmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsrep1.df.3,repeats=10,folds=5)
# rfeglmnet_results1.3[["avgScaledCoeff"]]


###Try a Ridge Regression just to see what it does with the variables:####
model.ridge <- glmnet(as.matrix(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")]),as.vector(train_data_scaled$LPS.Added),family="binomial",alpha = 0)
coef.ridge <- as.data.frame(as.matrix(coef(model.ridge, s=0.2))[-1,]) #doesn't seem to matter what we put as "s" when it is a ridge optimization
colnames(coef.ridge) <- "coefs"
coef.ridge$vars <- rownames(coef.ridge)
coef.ridge.nz <-  coef.ridge[which(coef.ridge$coefs !=0),]
coef.ridge.nz.ordered <- as.data.frame(coef.ridge.nz[order(abs(coef.ridge.nz$coefs),decreasing = TRUE),])

confusionMatrix(as.factor(round(predict(model.ridge,as.matrix(test_data_scaled[,-which(colnames(test_data_scaled)=="LPS.Added")]),s=0.2,type="response"),0)),as.factor(test_data_scaled$LPS.Added))

saveRDS(coef.ridge.nz.ordered,"GLMNET_RidgeRegression_Importance.rds")

### Custom glmnet function using ridge regression (alpha = 0) so that it keeps all provided predictors: ####
rfeglmnet <- function(x,y, testx=NULL,testy=NULL,repeats = 50, folds = 5,sizes = c(ncol(x):2),foldsreps = foldsreps1.df,lambda.set=0.3){
  sizes <- sizes[order(sizes, decreasing = TRUE)]
  modelinfo <- NULL
  #The function only works if the first size is the full size. Checking that 84 is one of the selected sizes
  check84 <- which(sizes == 84)
  if(length(check84)==0){
    sizes <- c(84,sizes)#Add a "full size" model if not inputted
  }
  
  varimp <- vector(mode='list', length=repeats)
  models <- vector(mode='list', length=repeats)
  
    #structuring input data (because some of the algorithms take x and y as part of a dataframe, others as two separate arguments)
  data <- x
  data$y <- y
  
   #If testing data is given, structuring test data (because some of the algorithms take x and y as part of a dataframe, others as two separate arguments)
  if(!is.null(testx) || !is.null(testy)){
    testing <- testx
    testing$y <- testy
  }
  
  FoldAccuracies <- data.frame(matrix(ncol=length(sizes),nrow=repeats*folds))
  colnames(FoldAccuracies) <- sizes
  allcoefs<- data.frame("var"=colnames(x))
  for(k in 1:repeats){
    varimp[[k]] <- vector(mode='list', length=folds)
    models[[k]] <- vector(mode='list', length=folds)
    kcoefs <- data.frame("var"=colnames(x))
    for(j in 1:folds){
        varimp[[k]][[j]] <- vector(mode='list', length=length(sizes))
        models[[k]][[j]] <- vector(mode='list', length=length(sizes))
        #itrain_index <- createDataPartition(data$y,p=.9,list=F) 
        jfoldindex <- foldsreps[,(k-1)*5 + j] #Folds/Reps are ordered by Reps and then by folds.
        jtrain_data <- data[jfoldindex,]
        jtest_data <- data[-jfoldindex,]
        predictors <- colnames(x)
        
      newentry <- data.frame(matrix(ncol=3,nrow=length(sizes)))
      colnames(newentry) <- c("Variables","Accuracy","NonZeroCoeffs")
      jcoefs <- data.frame("var"=colnames(x))
      
      for(i in 1:length(sizes)){
      size <- sizes[i]
      newentry$Variables[i] <- size
        
        #Generate a new model at the new size
        ipredictors <- jtrain_data[,c(which(colnames(jtrain_data) %in% predictors[1:size]))]
        imodel <- glmnet(x = as.matrix(ipredictors),y =as.vector(jtrain_data$y),family = "binomial",alpha = 0) #Ridge regression keeps all the coefficients provided
        
        #If no test set is given, check on the unused fraction of the training set:
      if(is.null(testx) | is.null(testy)){
        itestpredictors <- jtest_data[,which(colnames(jtest_data) %in% predictors[1:size])]
        check <- confusionMatrix(as.factor(round(predict(imodel,newx=as.matrix(itestpredictors),type="response",s=lambda.set),0)),as.factor(jtest_data$y))
      }else{
        #Check new model on provided test set
        itestpredictors <- testing[,which(colnames(testing) %in% predictors[1:size])]
        check <- confusionMatrix(as.factor(round(predict(imodel,newx=as.matrix(itestpredictors),type="response",s=lambda.set),0)),as.factor(testing$y))
      }
      
      newentry$Accuracy[i] <- check$overall["Accuracy"]
      
      coefs <- as.data.frame(as.matrix((coef(imodel,s=lambda.set))))
      coefs2 <- coefs[rownames(coefs) != "(Intercept)",]
      names(coefs2) <- rownames(coefs)[rownames(coefs)!= "(Intercept)"]
      coefs2[is.na(coefs2)] <- 0
      vimp <- data.frame( unname(coefs2), var = names(coefs2))
      colnames(vimp)[1]<-paste0("Vars",size,".",k,".",j)
      rownames(vimp) <- names(coefs2)
      vimp <- vimp[order(abs(vimp[,1]), decreasing = TRUE), , drop = FALSE]
      vimpscaled <- vimp
      vimpscaled[,1] <- as.numeric(abs(vimpscaled[,1])/max(abs(vimpscaled[,1])))
      iscaled0s <- which(vimpscaled[,1] == 0)
      if(length(iscaled0s) > 1)      reorder0 <- sample(iscaled0s) else reorder0 <- iscaled0s
      vimpscaled.reordered <- vimpscaled[c(which(vimpscaled[,1] != 0),reorder0),]
      #vimp
      
      #Save variable importance and model structure for each model
      varimp[[k]][[j]][[i]] <- vimp
      models[[k]][[j]][[i]] <- imodel
       
      FoldAccuracies[(k-1)*folds+j,i] <- check$overall["Accuracy"]
      
      newentry$NonZeroCoeffs[i] <- length(which(vimp[,1] != 0))
      
      joined <- left_join(jcoefs,vimpscaled.reordered,by="var")
      jcoefs <- joined
      coeforder <- vimp
      predictors <- vimpscaled.reordered$var
      }
      newkcoefs <- left_join(kcoefs,jcoefs,by="var")
      kcoefs <- newkcoefs
      
      modelinfo <- rbind(modelinfo,newentry)
    }
    newallcoefs <- left_join(allcoefs,kcoefs,by="var")
    allcoefs <- newallcoefs
  }
AvgAcc <- MeanSD(FoldAccuracies)
aveacc2 <- data.frame("Ave" = AvgAcc[1:(length(AvgAcc)/2)],"SD" =AvgAcc[(length(AvgAcc)/2 + 1):length(AvgAcc)] )

newallcoefs[is.na(newallcoefs)] = 0
rownames(newallcoefs)<- newallcoefs[,1]
avgCoeff <- data.frame("name" = newallcoefs[,1],"ave"=rowSums(newallcoefs[,-1])/(ncol(newallcoefs)-1),"sd" = apply(newallcoefs[,-1],1,sd))
avgCoeff.ordered <- avgCoeff[order(avgCoeff$ave,decreasing = TRUE),]

  sizemods.df <- data.frame(matrix(nrow=length(sizes),ncol=3))
  colnames(sizemods.df) <- c("Size","AccAve","AccSD")
  for(i in 1:length(sizes)){
    isizeinfo <- modelinfo[which(modelinfo$Variables == sizes[i]),]
    AvgInfo <- MeanSD(isizeinfo)
    aveinfo2 <- data.frame("Ave" = AvgInfo[1:(length(AvgInfo)/2)],"SD" =AvgInfo[(length(AvgInfo)/2 + 1):length(AvgInfo)])
    sizemods.df$Size[i] <- sizes[i]
    sizemods.df$AccAve[i] <- aveinfo2$Ave[which(rownames(aveinfo2)=="Accuracy")]
    sizemods.df$AccSD[i] <- aveinfo2$SD[which(rownames(aveinfo2)=="Accuracy")]
  }
  
  return(list("modelinfo"=modelinfo,"AveAcc"=aveacc2,"avgScaledCoeff"=avgCoeff.ordered,"AllAcc"=FoldAccuracies, "AllScaledCoeff"=newallcoefs,"varimp"=varimp,"models"=models,"SizeResults.df"=sizemods.df))
}

rfeglmnet_results1 <- rfeglmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsreps1.df,repeats=50,folds=5,lambda.set = 0.01) #very slight changes to importance order if lambda.set is 0.01 instead of 0.3
rfeglmnet_results1[["avgScaledCoeff"]] 
glmnetrfe_summary<-rfeglmnet_results1[["modelinfo"]]
glmnetrfe_size_vs_accuracy <- as.data.frame(rfeglmnet_results1[["SizeResults.df"]])

ggplot(glmnetrfe_size_vs_accuracy,aes(`Size`,`AccAve`))+geom_point()+theme_bw()

#saveRDS(rfeglmnet_results1,"GLMNET_RidgeRegRFE.rds")

rfeglmnet_results2 <- rfeglmnet(train_data_scaled[,-which(colnames(train_data_scaled)=="LPS.Added")],train_data_scaled$LPS.Added,foldsreps = foldsreps1.df,repeats=50,folds=5,lambda.set = 0.44) #very slight changes to importance order if lambda.set is 0.01 instead of 0.3
rfeglmnet_results2[["avgScaledCoeff"]] 
glmnetrfe_summary2<-rfeglmnet_results2[["modelinfo"]]
glmnetrfe_size_vs_accuracy2 <- as.data.frame(rfeglmnet_results2[["SizeResults.df"]])

ggplot(glmnetrfe_size_vs_accuracy2,aes(`Size`,`AccAve`))+geom_point()+theme_bw()
write_csv(glmnetrfe_size_vs_accuracy2,"GLMNet_RidgeRegRFE_l0p44.csv")

#                name         ave         sd
# CCL5           CCL5 0.980666798 0.06731634
# CXCL2         CXCL2 0.825035837 0.11580637
# LTB             LTB 0.806412425 0.13808492
# IL6             IL6 0.601491429 0.17270337
# CXCL1         CXCL1 0.587931892 0.12450228
# CCL20         CCL20 0.536497051 0.18584535
# CCL21         CCL21 0.487672893 0.20187060
# CCL2           CCL2 0.482129419 0.17247521
# CXCL8         CXCL8 0.475117692 0.21509475
# CSF3           CSF3 0.401507013 0.18146639
# BMP4           BMP4 0.396623148 0.15869547
# CCL17         CCL17 0.393866730 0.19760995
# CXCL5         CXCL5 0.385450630 0.16741352
# IL1B           IL1B 0.381721429 0.14798688
# TNFSF13B   TNFSF13B 0.341167364 0.19159132
# NODAL         NODAL 0.287422772 0.17888397
# IL15           IL15 0.278472673 0.14973992
# C5               C5 0.267318798 0.15842861
# TNF             TNF 0.210267323 0.16391023
# BMP7           BMP7 0.198801880 0.13799729
# CNTF           CNTF 0.198474886 0.12267220
# TNFSF11     TNFSF11 0.180638833 0.13925954
# VEGFA         VEGFA 0.171485147 0.10504532
# LTA             LTA 0.157349630 0.11308440
# MIF             MIF 0.155825624 0.11502461
# ADIPOQ       ADIPOQ 0.140082749 0.13961412
# IL18           IL18 0.135645046 0.09616216
# SPP1           SPP1 0.135610089 0.10906693
# TNFRSF11B TNFRSF11B 0.132349442 0.13626678
# CXCL16       CXCL16 0.127412280 0.11241598
# TNFSF10     TNFSF10 0.121040452 0.10493180
# CXCL11       CXCL11 0.119675208 0.11596482
# CXCL9         CXCL9 0.110735370 0.09861429
# IL22           IL22 0.106990843 0.09119963
# CXCL13       CXCL13 0.102737089 0.09607115
# IL7             IL7 0.100090739 0.09451698
# CXCL10       CXCL10 0.100010034 0.11147645
# IL16           IL16 0.097905248 0.08289395
# XCL1           XCL1 0.091207882 0.10163343
# CCL22         CCL22 0.087054316 0.08929288
# IL24           IL24 0.086151951 0.09637506
# IL9             IL9 0.083138325 0.09399841
# PPBP           PPBP 0.079759979 0.09567604
# IL12A         IL12A 0.069526519 0.09173412
# IL4             IL4 0.065928323 0.09002017
# IL1RN         IL1RN 0.064919465 0.07068560
# IL12B         IL12B 0.056865298 0.10000462
# CCL19         CCL19 0.056533381 0.08115927
# IL11           IL11 0.056207361 0.06995900
# CSF1           CSF1 0.055581235 0.07877257
# THPO           THPO 0.054990362 0.07902933
# GPI             GPI 0.052735840 0.07031547
# IL21           IL21 0.050765549 0.06596560
# CX3CL1       CX3CL1 0.048906417 0.06006283
# IFNA2         IFNA2 0.047106585 0.07134413
# CXCL12       CXCL12 0.041135762 0.07477614
# TGFB2         TGFB2 0.040759205 0.05913769
# IFNG           IFNG 0.035536729 0.05067322
# MSTN           MSTN 0.035102765 0.07027167
# IL3             IL3 0.034118412 0.05040967
# CCL24         CCL24 0.031464194 0.06085994
# CCL3           CCL3 0.030268565 0.04887838
# CCL1           CCL1 0.027923201 0.04664003
# IL13           IL13 0.025744120 0.04432153
# CCL7           CCL7 0.025189055 0.05347134
# CCL13         CCL13 0.023041455 0.04205392
# IL27           IL27 0.021928750 0.04012295
# IL17A         IL17A 0.020479247 0.03995100
# IL23A         IL23A 0.019535577 0.04492781
# BMP6           BMP6 0.018795195 0.03999925
# IL1A           IL1A 0.018592467 0.04268904
# OSM             OSM 0.018582223 0.04183039
# BMP2           BMP2 0.015437089 0.03882935
# CD40LG       CD40LG 0.013257703 0.02988507
# CSF2           CSF2 0.013103100 0.03195263
# IL5             IL5 0.012464529 0.03185979
# FASLG         FASLG 0.012025556 0.02906201
# LIF             LIF 0.011868343 0.02832661
# IL17F         IL17F 0.010503397 0.03553444
# CCL11         CCL11 0.008055442 0.02341362
# IL10           IL10 0.008022650 0.02364066
# CCL18         CCL18 0.007820709 0.02228306
# IL2             IL2 0.006136775 0.01928785
# CCL8           CCL8 0.004054732 0.01398808



```
##Tune SVM
```{r SVM basic tune}
#svmgrid=expand.grid(C = c(5,1,5e-1,1e-1,5e-2,1e-2,5e-3,1e-3,5e-4,1e-4,5e-5,1e-5))
svmgrid=expand.grid(C=sort(c(2:10 %o% 10^(1:-5),0.00001),decreasing = TRUE))
#Unscaled data:####
svm.pre.unscaled.mod <- train(as.factor(LPS.Added)~.,data=train_data,method = "svmLinear",
                    trControl = trainControl(index = foldsreps1),tuneGrid=svmgrid)
svm.pre.unscaled.mod$bestTune
#        C
# 27 0.009

      SVM_UNscaled_foldsreps_plot <-ggplot(data = svm.pre.unscaled.mod$results,mapping=aes(x=C,y=Accuracy))+
        geom_line()+
        scale_x_log10(breaks=c(100,10,1,1e-1,1e-2,1e-3,1e-4,1e-5))+
        #xlim(Alpha.factors.ordered)+
        ylab("SVM Accuracy")+xlab("SVM `cost` values")+
        theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10,angle = 45,vjust = 0.5))+
        theme(axis.text.y = element_text(size = 10))
        #+scale_x_continuous(breaks=c(1,3,5,7,9,11))

#Scaled data:####

svm.pre.scaled.mod <- train(as.factor(LPS.Added)~.,data=train_data_scaled,method = "svmLinear",
                    trControl = trainControl(index = foldsreps1),tuneGrid=svmgrid)
svm.pre.scaled.mod$bestTune 
#        C
# 27 0.009

      SVM_scaled_foldsreps_plot <-ggplot(data = svm.pre.scaled.mod$results,mapping=aes(x=C,y=Accuracy))+
        geom_line()+
        scale_x_log10(breaks=c(5,1,1e-1,1e-2,1e-3,1e-4,1e-5))+
        #xlim(Alpha.factors.ordered)+
        ylab("SVM Accuracy")+xlab("SVM `cost` values")+
        theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10,angle = 45,vjust = 0.5))+
        theme(axis.text.y = element_text(size = 10))

```
##Tune NB
```{NB basic tune}
nb_grid=expand.grid(laplace=seq(from=0,to=10,by=0.5),
                    adjust = 1,
                    usekernel = c(TRUE,FALSE))
laplace.factors.ordered <- levels(as.factor(nb_grid$laplace))[levels(as.factor(nb_grid$laplace)) %>% as.numeric %>% order(decreasing=FALSE)]

#Predefined folds/reps with UNSCALED data:####
nb.pre.mod.UNscaled <- train(x=train_data[,-1],y=as.factor(train_data$LPS.Added),method = "naive_bayes",trControl = trainControl(index = foldsreps1),tuneGrid=nb_grid)
nb.pre.mod.UNscaled$bestTune
#   laplace usekernel adjust
# 2       0      TRUE      1
saveRDS(nb.pre.mod.UNscaled,"NB_UNSCALED_smallgrid.rds")
nb.pre.mod.UNscaled <- readRDS("NB_UNSCALED_smallgrid.rds")
check <- nb.pre.mod.UNscaled$results
  
nb.pre.mod.UNscaled3 <- train(x=train_data[,-1],y=as.factor(train_data$LPS.Added),method = "naive_bayes",trControl = trainControl(index = foldsreps1),tuneGrid=nb_grid3)
nb.pre.mod.UNscaled3$bestTune

nb.pre.mod.UNscaled4 <- train(x=train_data[,-1],y=as.factor(train_data$LPS.Added),method = "naive_bayes",trControl = trainControl(index = foldsreps1),tuneGrid=nb_grid4)
nb.pre.mod.UNscaled4$bestTune
saveRDS(nb.pre.mod.UNscaled4,"NB_UNSCALED_smallgrid_adjustop.rds")
nb.pre.mod.UNscaled4 <- readRDS("NB_UNSCALED_smallgrid_adjustop.rds")

    #plot UNSCALED####
    NB_UNscaled_foldsreps_plot <-ggplot(data = nb.pre.mod.UNscaled$results,mapping=aes(x=usekernel,y=as.factor(laplace),fill=Accuracy))+
        geom_tile()+scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks,to=c(0,1)), limits=c(0.5,1))+
        ylim(laplace.factors.ordered)+
        #xlim(Alpha.factors.ordered)+
        #ggtitle("Validation Set")+#ggtitle("GLMNET Scaled Data - Manual Opt\n Validation Accuracy")+
        ylab("NB `laplace` smoothing values")+xlab("NB `usekernel`")+
        theme(axis.text=element_text(size=12),axis.title.y=element_text(size=14,face="bold"))+
                theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))#+scale_x_continuous(breaks=c(1,3,5,7,9,11))

        ggplot(data = nb.pre.mod.UNscaled4$results, aes(x=as.factor(usekernel),y=factor(adjust)))+geom_tile(aes(fill=Accuracy))+theme_classic2()+
          xlab("usekernel")+ylab("adjust")+
  # facet_grid(datatype~.)+
  scale_fill_viridis(direction=-1,option = "D")
        
#Predefined folds/reps with SCALED data:####
nb.pre.mod.scaled <- train(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),method = "naive_bayes",trControl = trainControl(index = foldsreps1),tuneGrid=nb_grid)
nb.pre.mod.scaled$bestTune
#   laplace usekernel adjust
# 2       0      TRUE      1
saveRDS(nb.pre.mod.scaled,"NB_SCALED_smallgrid.rds")
nb.pre.mod.scaled <- readRDS("NB_SCALED_smallgrid.rds")
check2 <- nb.pre.mod.scaled$results

nb.pre.mod.scaled4 <- train(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),method = "naive_bayes",trControl = trainControl(index = foldsreps1),tuneGrid=nb_grid4)
nb.pre.mod.scaled4$bestTune
saveRDS(nb.pre.mod.scaled4,"NB_SCALED_smallgrid_adjustop.rds")
nb.pre.mod.scaled4 <- readRDS("NB_SCALED_smallgrid_adjustop.rds")

    #plot SCALED####
    NB_scaled_foldsreps_plot <-ggplot(data = nb.pre.mod.scaled$results,mapping=aes(x=usekernel,y=as.factor(laplace),fill=Accuracy))+
        geom_tile()+scale_fill_gradientn(colours = c("red","white","blue"),values = rescale(Acc_breaks,to=c(0,1)), limits=c(0.5,1))+
        ylim(laplace.factors.ordered)+
        #xlim(Alpha.factors.ordered)+
        #ggtitle("Validation Set")+#ggtitle("GLMNET Scaled Data - Manual Opt\n Validation Accuracy")+
        ylab("NB `laplace` smoothing values")+xlab("NB `usekernel`")+
        theme(axis.text=element_text(size=12),axis.title.y=element_text(size=14,face="bold"))+
                theme_classic()+
        theme(axis.title.y=element_text(size=14,face="bold"))+
        theme(axis.title.x=element_text(size=14,face="bold"))+
      theme(axis.text.x = element_text(size = 10))+
        theme(axis.text.y = element_text(size = 10))#+scale_x_continuous(breaks=c(1,3,5,7,9,11))

    ggplot(data = nb.pre.mod.scaled4$results[which(nb.pre.mod.scaled4$results$usekernel == "TRUE"),], aes(x=as.factor(adjust),y=factor(laplace)))+geom_tile(aes(fill=Accuracy))+
  # facet_grid(datatype~.)+
  scale_fill_viridis(direction=-1,option = "D")
    
        ggplot(data = nb.pre.mod.scaled4$results, aes(x=as.factor(usekernel),y=factor(adjust)))+geom_tile(aes(fill=Accuracy))+theme_classic2()+
          xlab("usekernel")+ylab("adjust")+
  # facet_grid(datatype~.)+
  scale_fill_viridis(direction=-1,option = "D")
    

```
##Heatmaps of hyperparameters
```{Heatmaps of hyperparameters}

Scaled_v_Unscaled_fig<-ggarrange(NN_scaled_foldsreps_hm+ rremove("xlab")+ggtitle("NN"),
                                 RF_scaled_foldsreps_hm+ rremove("xlab")+ggtitle("RF"),
                                 GLMNet_scaled_foldsreps_hm3+ rremove("xlab")+ggtitle("GLM"),
                                 NB_scaled_foldsreps_plot+ rremove("xlab")+ggtitle("NB"),
                                 SVM_scaled_foldsreps_plot+ rremove("xlab")+ggtitle("SVM"),
                                 NN_unscaled_foldsreps_hm+ggtitle(""),
                                 RF_unscaled_foldsreps_hm+ggtitle(""),
                                 GLMNet_unscaled_foldsreps_hm3+ggtitle(""),
                                 NB_UNscaled_foldsreps_plot+ggtitle(""),
                                 SVM_UNscaled_foldsreps_plot+ggtitle(""),
                                     ncol=5,nrow=2,common.legend=TRUE,labels="auto",legend="right")

GLMNET_Scaled_cv.test.val<-annotate_figure(GLMNET_Scaled_cv.test.val,bottom = text_grob("GLM `\u03b1` values", size=14,face="bold",vjust = 0.1),left = text_grob("GLM \u03bb values", size=14,face="bold",rot = 90))
```


# Recursive Feature Selection

After deciding on a train/test split of **XXX/XXX** based on **....** We can set 
the seed (so our training data and models are the same each time). Then we can 
run recursive feature selection to optimize hyper parameters and build smaller 
models.

This will be set up to optimize the following parameters for each model:

-   RF: mtry across a range of (??-??). the number of trees likely does not need
to be tested since largely 500 trees has been sufficient

-   SVM: cost across a range of (??-??)

-   GLM: lambda same as in the earlier example

-   NB: none!

-   NN: number of hidden layers (1,2,3); and number of nodes in each layer (??); 
not going to mess with defaults of learning rates.


We will perform RFE on the 5 models using 10 repetitions of 10 fold cross 
validation. 

```{r RFE set up}
library(doParallel)
cl <- makeCluster(8)
registerDoParallel(cl)
ctrl_cvrep <- rfeControl(method = "repeatedcv",number=2,repeats = 3,allowParallel = TRUE,saveDetails = TRUE)
sizes <- c(2,4,6,8)
run_rfe <- function(ctrl,x,y,sizes,varimp="default",...){
  if(varimp=="default"){
    result <- rfe(x,y,sizes,rfeControl=ctrl,...)
  }else{
    ctrl$functions$rank <- nbFuncs$rank
    result <- rfe(x,y,sizes,rfeControl=ctrl,...)
  }
  return(result)
}
glm_ctrl <-ctrl_cvrep
glm_ctrl$functions <- lrFuncs
glm_ctrl$functions$rank <- nbFuncs$rank ## applies a filter approach to variable imp
glm_grid=expand.grid(lambda=c(1e-1,1e-2,1e-3,1e-4,1e-5),
                     alpha=1)
rfe_glm <- run_rfe(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),
                   sizes = sizes,ctrl=glm_ctrl,
                   method="glmnet",family="binomial",tuneGrid=glm_grid)

rf_ctrl <- ctrl_cvrep
rf_ctrl$functions <- rfFuncs
# rf_ctrl$functions$rank <- nbFuncs$rank ## applies a filter approach to variable imp
rf_grid=expand.grid(mtry=c(1:25))
rfe_rf <- run_rfe(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),
                   sizes = sizes,ctrl=rf_ctrl,
                   method="rf",tuneGrid=rf_grid)
# 
nb_ctrl <- ctrl_cvrep
nb_ctrl$functions <- nbFuncs
nb_grid=expand.grid(laplace=seq(from=0,to=1,by=.1),
                    usekernel = c(TRUE,FALSE))
rfe_nb <- run_rfe(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),
                   sizes = sizes,ctrl=nb_ctrl,tuneGrid=nb_grid)
# 
svm_ctrl <- ctrl_cvrep
svm_ctrl$functions <- caretFuncs
svm_ctrl$functions$rank <- nbFuncs$rank ## applies a filter approach to variable imp
svm_grid <- expand.grid(C=c(5,1,5e-1,1e-1,5e-2,1e-2,5e-3,1e-3,5e-4,1e-4,5e-5,1e-5))
rfe_svm <- run_rfe(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),
                   sizes = sizes,ctrl=svm_ctrl,
                  method = "svmLinear",
                  tuneGrid=svm_grid)

nn_ctrl <-ctrl_cvrep
nn_ctrl$functions <- caretFuncs
nn_grid <-  expand.grid(size=c(12:13),
                     decay=c(1e-3,1e-4))
rfe_nn <- run_rfe(x=train_data_scaled[,-1],y=as.factor(train_data_scaled$LPS.Added),
                   sizes = sizes,ctrl=nn_ctrl,
                  method = "nnet",
                  tuneGrid=nn_grid,trace=FALSE,linout=FALSE)


```

```{r RFE Analysis}
library(ggthemes)

rfe_results <- list(rfe_rf,rfe_glm,rfe_nb,rfe_svm,rfe_nn)
names(rfe_results) <- c("RF","GLM","NB","SVM","NN")

rfe_df <- data.frame()
for(i in 1:length(rfe_results)){
  tmp_df <- rfe_results[[i]]$results %>% 
    mutate(model=names(rfe_results)[i]) %>% 
    mutate(isBestSubset=ifelse(Variables==rfe_results[[i]]$bestSubset,TRUE,FALSE)) %>% 
    mutate(optVars=ifelse(Variables==rfe_results[[i]]$bestSubset,
                          paste0(rfe_results[[i]]$optVariables,collapse = ","),NA)) 
  rfe_df <- rfe_df %>% bind_rows(tmp_df)
}
rfe_df <- rfe_df %>% 
         mutate(AccSDHigh=ifelse(Accuracy+AccuracySD>1,1,Accuracy+AccuracySD)) %>% 
  distinct()

ggplot(rfe_df %>% filter(model!="GLM"),aes(x=Variables,y=Accuracy,color=model,fill=model)) +
  geom_line(size=.75)+
  geom_point(data=rfe_df %>% filter(isBestSubset)%>% filter(model!="GLM"),pch=19,size=3,inherit.aes = TRUE)+
  geom_ribbon(aes(ymax = AccSDHigh, ymin = Accuracy - AccuracySD),linetype=2, alpha = 0.25)+
  facet_wrap(.~model,nrow=1)+
  theme_few()+ 
  scale_fill_few() +
  scale_color_few()+
  # scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        panel.grid.major.y = element_line(color = "grey80",linetype = 3),
        legend.position = "none",
        )
  # coord_cartesian(ylim=c(0,.5),xlim=c(.1,.9))

rfe_df %>% filter(isBestSubset) %>% select(model,Variables,Accuracy,optVars)
 print(xtable(rfe_df %>% filter(isBestSubset) %>% select(model,Variables,Accuracy,optVars),digits = 6),include.rownames=FALSE)

```

## Variable importance

```{RFE Var Imp}
gene.list.fill <- LPS.qPCR.norm1.tidy %>% select(Gene.Symbol) %>% 
                mutate( Gene.Symbol=ifelse(Gene.Symbol=="1L13","IL13",Gene.Symbol))%>% 
                filter(!(Gene.Symbol%in%c("ACTB","B2M","GAPDH","HPRT1","RPLP0","HGDC","RTC","PPC"))) %>% 
                distinct()

rfe_var_imp <- data.frame()
for(i in 1:length(rfe_results)){
  tmp_df <- rfe_results[[i]]$variables%>%
    mutate(model=names(rfe_results)[i])%>%
    select(model,Overall,Resample,var) %>%
    distinct() %>% 
    full_join(gene.list.fill ,
              by=c("var"="Gene.Symbol")) %>%
    mutate(Overall=ifelse(is.na(Overall),0,Overall)) %>% 
    mutate(optVars=ifelse(var %in% rfe_results[[i]]$optVariables,TRUE,FALSE)) %>%
    separate(Resample, into=c("Fold","Rep"),remove = FALSE) %>%
    group_by(Fold,Rep) %>%
    
    arrange(desc(Overall)) %>%
    mutate(rank=ifelse(Overall==0, 84, row_number())) %>%
    mutate(rankNA=ifelse(Overall==0, NA, row_number()))%>%
    mutate(Scaled_imp=(Overall-min(Overall))/(max(Overall)-min(Overall))) %>% 
    ungroup()
  rfe_var_imp <- rfe_var_imp %>% bind_rows(tmp_df)
}
# TODO: make sure Kristen's code is above this 
glm_varimp <- read.csv("GLMNet_ScaledCoeffs_ap1_lp44.csv") %>% 
  select(-X) %>% 
  pivot_longer(contains("Lambda0.44."),names_to = "Resample",values_to = "Overall",names_prefix = "Lambda0.44.") %>% 
  group_by(Resample) %>% 
  arrange(desc(abs(Overall))) %>%
  mutate(rank=ifelse(Overall==0, 84, row_number())) %>%
  mutate(rankNA=ifelse(Overall==0, NA, row_number())) %>% 
  separate(Resample,into=c("Fold","Rep"),remove = FALSE) %>% 
  mutate(Fold=paste0("Fold",Fold),
         Rep=paste0("Rep",Rep)) %>% 
  mutate(absOverall=abs(Overall)) %>% 
  mutate(Scaled_imp=(absOverall-min(absOverall))/(max(absOverall)-min(absOverall))) %>% 
  ungroup() %>% 
  mutate(model="GLM_EN") 

rfe_var_imp <- bind_rows(rfe_var_imp,glm_varimp)%>% 
  mutate(model=factor(model,levels=c("RF","NN","SVM","NB","GLM","GLM_EN"),
                      labels=c("Random\n Forest","Neural\n Net","Support\n Vector\n Machine","Naive\n Bayes","GLM","GLM\n Elastic\n Net")))

top_vars <- rfe_var_imp %>%filter(model!="GLM") %>%  group_by(var) %>% 
  summarise(mean_rank=mean(rank)) %>% 
  arrange(mean_rank) %>% 
  mutate(var.rank=row_number()) #%>% 
  # mutate(var=factor(var,levels=var))

theme.set <- theme_few()+ 
  # scale_x_continuous(breaks = seq(0, 1, by = .2),minor_breaks = seq(0, 1, .1))+
  theme(text = element_text(size=30),
        axis.title = element_text(size=20),
        axis.text = element_text(size=15),
        panel.grid.major.x = element_line(color = "grey80",linetype = 3),
        panel.grid.major.y = element_line(color = "grey80",linetype = 3),
        legend.position = "none",
        )

show.top.n <- top_vars$var[1:8]
rfe_var_imp %>% filter(model!="GLM") %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%top_vars$var[1:10]) %>% 
  # filter(Overall>0) %>%
  # filter(optVars) %>%
  # filter(rank<20) %>%
  # mutate(Overall=ifelse(model=="GLM",log(Overall),Overall)) %>%
  ggplot(aes(x=var,y=Scaled_imp))+geom_violin()+
  facet_wrap(model~.,scales="free")+
  scale_fill_few() +
  scale_color_few()+
  theme.set#
# 
rfe_var_imp %>% filter(model!="GLM") %>%
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%top_vars$var[1:10]) %>% 
  filter(model!="GLM") %>%
  # filter(optVars) %>%
  # filter(rank<20) %>%
  # mutate(Overall=ifelse(model=="GLM",log(Overall),Overall)) %>%
  ggplot(aes(x=model,y=Scaled_imp,color=model,fill=model))+geom_violin(scale="width")+
  facet_wrap(var~.,nrow=2)+
  theme_minimal()+
  theme(text = element_text(size=16))+
  xlab(NULL)+
  ylab("Scaled Importance")+
  scale_fill_few() +
  scale_color_few()+
  theme.set
rfe_var_imp %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%top_vars$var[1:16]) %>% 
  # filter(Overall>0) %>%
  # filter(optVars) %>%
  # filter(rank<20) %>%
  # mutate(Overall=ifelse(model=="GLM",log(Overall),Overall)) %>%
  ggplot(aes(x=var,y=rankNA,color=model,fill=model))+geom_violin()+
  facet_wrap(var~.,scales="free_x")
summary_rfe_var_imp <- rfe_var_imp %>% 
  group_by(model,var) %>% 
  summarise(meanScaledImp=mean(Scaled_imp,na.rm=T),
            medianScaledImp=median(Scaled_imp,na.rm=T),
            meanRankNA=mean(rankNA,na.rm=T),
            medRankNA=median(rankNA,na.rm=T),
            meanRank=mean(rank,na.rm=T),
            medRank=median(rank,na.rm=T)) %>% 
  ungroup()

summary_rfe_var_imp %>% mutate(var=factor(var,levels=top_vars$var)) %>% filter(model!="GLM") %>% 
  ggplot( aes(model, var, fill= meanScaledImp)) + 
  geom_tile(color = "gray")+
    scale_fill_viridis(discrete=FALSE) +
  theme.set+
  xlab(NULL)+
  ylab(NULL)+
  theme(axis.text = element_text(size=12))
summary_rfe_var_imp %>% mutate(var=factor(var,levels=top_vars$var)) %>% 
  ggplot( aes(model, var, fill= meanRank)) + 
  geom_tile()+
    scale_fill_viridis(discrete=FALSE,direction = -1)
summary_rfe_var_imp %>% mutate(var=factor(var,levels=top_vars$var)) %>% 
  ggplot( aes(model, var, fill= medRank)) + 
  geom_tile()+
    scale_fill_viridis(discrete=FALSE,direction = -1)# 

library(ggridges)
library(ggplot2)

# Diamonds dataset is provided by R natively
#head(diamonds)
 
# basic example
rfe_var_imp %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%c(top_vars$var[1:6],top_vars$var[20:25])) %>% 
ggplot( aes(x = Scaled_imp, y = model, fill = model)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")+
  facet_wrap(.~var)+
  coord_cartesian(xlim=c(0,1))
rfe_var_imp %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%top_vars$var[1:12]) %>% 
ggplot( aes(x = Scaled_imp, y = model, fill = model)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")+
  facet_wrap(.~var)+
  coord_cartesian(xlim=c(0,1))

rfe_var_imp %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%top_vars$var[1:20]) %>% 
  filter(model!="GLM") %>% 
ggplot( aes(x = Scaled_imp, y = model, fill = model)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")+
  facet_wrap(.~var)+
  coord_cartesian(xlim=c(0,1))

rfe_var_imp %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% filter(var%in%top_vars$var[1:12]) %>% 
  filter(model!="GLM") %>% 
ggplot( aes(x = rank, y = model, fill = model)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")+
  facet_wrap(.~var)+
  coord_cartesian(xlim=c(0,20))



# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)
glm_varimp <-varImp(cv_glm,lambda = .44)%>%   
  rownames_to_column(var = "Gene.Symbol") %>%  
  full_join(gene.list.fill) %>% 
  mutate(Overall=ifelse(is.na(Overall),0,Overall)) %>% 
  mutate(Scaled_imp=(Overall-min(Overall))/(max(Overall)-min(Overall))) %>% 
  mutate(model="GLM_EN")%>% 
  rename(Imp=Overall)
svm_varimp <-varImp(rfe_svm2)%>%   
  rownames_to_column(var = "Gene.Symbol") %>%  
  full_join(gene.list.fill) %>% 
    mutate(Overall=ifelse(is.na(Overall),0,Overall)) %>% 
  mutate(Scaled_imp=Overall)%>% 
  mutate(model="SVM")%>% 
  rename(Imp=Overall)
nb_varimp <-varImp(rfe_nb)%>%   
  rownames_to_column(var = "Gene.Symbol") %>%  
  full_join(gene.list.fill)%>% 
  mutate(Overall=ifelse(is.na(Overall),0,Overall)) %>% 
  mutate(Scaled_imp=Overall)%>% 
  mutate(model="NB")%>% 
  rename(Imp=Overall)
rf_varimp <- varImp(rfe_rf$fit) %>%   
  rownames_to_column(var = "Gene.Symbol") %>%  
  full_join(gene.list.fill)%>% 
  rename(varimp_raw="0") %>% 
    mutate(varimp_raw=ifelse(is.na(varimp_raw),0,varimp_raw)) %>% 
  mutate(Scaled_imp=(varimp_raw-min(varimp_raw))/(max(varimp_raw)-min(varimp_raw)))%>% 
  # rownames_to_column(var = "Gene.Symbol") %>% 
  mutate(model="RF") %>% 
  rename(Imp=varimp_raw)
nn_varimp <-varImp(rfe_nn$fit)$importance %>%   
  rownames_to_column(var = "Gene.Symbol") %>%  
  full_join(gene.list.fill)%>% 
    mutate(Overall=ifelse(is.na(Overall),0,Overall)) %>% 
  mutate(Scaled_imp=Overall/100)%>% 
  # rownames_to_column(var = "Gene.Symbol") %>% 
  mutate(model="NN") %>% 
  rename(Imp=Overall)

# library(ggbump)

all_varimp <- bind_rows(nn_varimp,rf_varimp,nb_varimp,svm_varimp,glm_varimp) %>% 
  select(model,Gene.Symbol,Imp,Scaled_imp) %>% 
  group_by(model) %>% 
  arrange(desc(Scaled_imp)) %>% 
  mutate(rank=ifelse(Scaled_imp==0,84,row_number()),
         TopN=ifelse(rank<=15,1,0)) %>% 
  ungroup() %>%
  group_by(Gene.Symbol) %>% 
  mutate(TopN_any=max(TopN)) %>% 
  ungroup()%>% 
  mutate(model=factor(model,levels=c("RF","NN","SVM","NB","GLM","GLM_EN"),
                      labels=c("Random\n Forest","Neural\n Net","Support Vector\n Machine","Naive\n Bayes","GLM","GLM\n Elastic Net")))


plot_varimp1 <- all_varimp %>% filter(TopN==1) %>% 
  group_by(Gene.Symbol) %>% 
  count() %>% 
  arrange(desc(n))
plot_varimp1 %>% 
  mutate(Gene.Symbol=factor(Gene.Symbol,levels=plot_varimp1$Gene.Symbol)) %>% 
  ggplot(aes(x=Gene.Symbol,y=n)) +
  geom_bar(stat="identity")

plot_varimp12 <- all_varimp %>% filter(TopN==1) 
plot_varimp12 %>% 
  mutate(Gene.Symbol=factor(Gene.Symbol,levels=rev(plot_varimp1$Gene.Symbol))) %>% 
  ggplot(aes(x=Gene.Symbol,y=TopN,color=model,fill=model)) +
  geom_bar(stat="identity",position = "stack")+
  theme.set+
  theme(legend.position = c(.8, 0.2),
        legend.text = element_text(size=14))+
  scale_fill_few()+
  scale_color_few()+
  ylab("Number of Models")+
  xlab(NULL)+
  coord_flip()

show.top.n <- 15
all_varimp%>% mutate(Gene.Symbol=factor(Gene.Symbol,levels=top_vars$var))  %>% filter(TopN==1) %>% 
  ggplot(aes(x=Gene.Symbol,y=Scaled_imp,color=model,fill=model))+
  geom_bar(stat = 'identity')+
  facet_wrap(.~model,nrow = 1)+
  coord_flip()


all_varimp %>% mutate(Gene.Symbol=factor(Gene.Symbol,levels=top_vars$var)) %>% filter(Gene.Symbol%in%top_vars$var[1:16]) %>% 
  ggplot(aes(x=Gene.Symbol,y=Scaled_imp,color=model,fill=model))+
  geom_bar(stat = 'identity',position="dodge")+
  facet_wrap(.~Gene.Symbol,scales="free_x")

all_varimp %>% mutate(Gene.Symbol=factor(Gene.Symbol,levels=top_vars$var)) %>% filter(Gene.Symbol%in%top_vars$var[1:16]) %>% 
  ggplot(aes(x=Gene.Symbol,y=Scaled_imp,color=model,fill=model))+
  geom_bar(stat = 'identity',position="dodge")#+
  # facet_wrap(.~Gene.Symbol,scales="free_x")
all_varimp %>% mutate(Gene.Symbol=factor(Gene.Symbol,levels=top_vars$var))  %>% filter(TopN_any==1) %>% 
  ggplot(aes(x=Gene.Symbol,y=Scaled_imp,color=model,fill=model))+
  geom_bar(stat = 'identity',position="dodge")
all_varimp %>% #mutate(Gene.Symbol=factor(Gene.Symbol,levels=top_vars$var)) %>% 
  ggplot( aes(model, Gene.Symbol, fill= Scaled_imp)) + 
  geom_tile()+
    scale_fill_viridis(discrete=FALSE) 
all_varimp %>% #mutate(Gene.Symbol=factor(Gene.Symbol,levels=top_vars$var)) %>% 
  ggplot( aes(model, Gene.Symbol, fill= rank)) + 
  geom_tile()+
    scale_fill_viridis(discrete=FALSE,direction = -1)   # coord_flip()

wide_varimp <- all_varimp %>% filter(rank<=10)%>% select(model,Gene.Symbol,Imp) %>% 
  pivot_wider(names_from = "model",values_from = "Imp",values_fill = 0)




# Libraries
library(hrbrthemes)
library(GGally)
library(viridis)

# Data set is provided by R natively
# data <- iris
# Plot
ggparcoord(wide_varimp ,
    columns = 2:6, groupColumn = 1,scale="uniminmax",
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for the Iris Data",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )

wide_varimp_all <- rfe_var_imp %>% pivot_wider(names_from = "model",values_from = "Overall",values_fill = 0)

ggparcoord(wide_varimp_all %>% filter(var%in%c("CCL5","CXCL1","IL6","CCL20","TNF")),
    columns = 7:11, groupColumn = 4,scale="uniminmax",
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for the Iris Data",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )


wide_varimp_all_rank <- rfe_var_imp %>% select(-Overall)%>% pivot_wider(names_from = "model",values_from = "rank",values_fill = 84) %>% 
  mutate(var=factor(var,levels=top_vars$var)) 



ggparcoord(wide_varimp_all_rank %>% filter(var%in%show.top.n),
    columns = 6:10, groupColumn = 4,scale="globalminmax",
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for the Iris Data",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )+
  scale_y_reverse(breaks = c(1:10,6:15*2,7:16*5,84))+
  coord_cartesian(ylim = c(84,1))+
  facet_wrap(.~var)
# load the library
# wide_varimp_all_rank <-

test <-  rfe_var_imp %>% select(-Overall)%>% pivot_wider(names_from = "model",values_from = "rank",values_fill = NA) %>% 
  mutate(var=factor(var,levels=top_vars$var))  %>% pivot_longer(6:10,names_to = "Model",values_to = "Rank")


ggplot(test %>% filter(var%in%show.top.n),
       aes(x=var,y=Rank,color=Model,fill=Model))+geom_violin()+facet_wrap(.~var,scales = "free") + theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )+
  coord_cartesian(ylim = c(84,1))


ggplot(test %>% filter(var%in%show.top.n),
       aes(x=var,y=Rank,color=var,fill=var))+geom_violin()+facet_wrap(.~Model,scales = "free") + theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )+
  coord_cartesian(ylim = c(84,1))

test <-  rfe_var_imp %>% pivot_wider(names_from = "model",values_from = "Overall",values_fill = NA)  %>% 
  pivot_longer(7:11,names_to = "Model",values_to = "Imp")


ggplot(test %>% filter(var%in%show.top.n),
       aes(x=var,y=Imp,color=Model,fill=Model))+geom_violin()+facet_wrap(.~var,scales = "free") + theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )


ggplot(test %>% filter(var%in%show.top.n),
       aes(x=var,y=Imp,color=var,fill=var))+geom_violin()+facet_wrap(.~Model,scales = "free") + theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )




library(forcats)




# Reorder following the value of another column:
all_varimp %>%
  group_by(model) %>% 
  arrange(desc(Imp)) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Gene.Symbol=factor(Gene.Symbol, levels=Gene.Symbol)) %>% 
  filter(rank<=15) %>% 
# mutate(Gene.Symbol = fct_reorder(as.factor(Gene.Symbol), rank)) %>%
  ggplot( aes(x=Gene.Symbol, y=Imp,color=model,fill=model)) +
    geom_bar(stat="identity",position="dodge", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()+
    facet_wrap(.~model,scales="free",nrow=1)
 
# Reverse side
data %>%
  mutate(name = fct_reorder(name, desc(val))) %>%
  ggplot( aes(x=name, y=val)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()


ggplot(all_varimp , aes(x = model, y = rank, color = Gene.Symbol,group=Gene.Symbol)) +
  geom_point()+
  geom_line()+
  scale_y_reverse(breaks = show.top.n:1)+
  coord_cartesian(ylim = c(show.top.n,1),
                  xlim = c(.70,5.3)) +
  theme(legend.position = "none") +
geom_text(data = all_varimp %>% filter(model == "GLM"),
          aes(label = Gene.Symbol, x = .75) , hjust = .85, fontface = "bold", color = "#888888", size = 4) +
  geom_text(data = all_varimp %>% filter(model == "SVM"),
            aes(label = Gene.Symbol, x = 5.25) , hjust = 0.15, fontface = "bold", color = "#888888", size = 4) 



ggplot(data = all_varimp, aes(x = model, y = Scaled_imp, group = Gene.Symbol)) +
  geom_line(aes(color = Gene.Symbol, alpha = 1), size = 2) +
  geom_point(aes(color = Gene.Symbol, alpha = 1), size = 4) +
  geom_point(color = "#FFFFFF", size = 1) +
  # scale_y_reverse(breaks = show.top.n:1) +
  scale_x_discrete(breaks = unique(all_varimp$model), expand = c(.25, .25)) +
  geom_text(data = all_varimp %>% filter(model == "GLM"),
          aes(label = Gene.Symbol, x = .75) , hjust = .85, fontface = "bold", color = "#888888", size = 4) +
  geom_text(data = all_varimp %>% filter(model == "SVM"),
            aes(label = Gene.Symbol, x = 5.25) , hjust = 0.15, fontface = "bold", color = "#888888", size = 4) +
  # coord_cartesian(ylim = c(show.top.n,1)) + 
  theme(legend.position = "none") 

```


# Final Models Error Analysis
```{r Errors, message=FALSE, warning=FALSE}
generate.errors.ML <- function(model.string,model.obj,predictors){
  y <- predictors$LPS.Added
  if(model.string %in% c("RF","NB","GLM")){
    observed <- as.factor(y)
    CM <- caret::confusionMatrix(predict(model.obj, predictors)$pred,observed)
  } else if(model.string%in%c("NN","SVM")){
    observed <- as.factor(y)  
    CM <- caret::confusionMatrix(predict(model.obj, predictors),observed)
  }
    # } else if(grepl("NN",model.string)){
    #   observed <- as.factor(y)
    #   CM <- caret::confusionMatrix(factor(predict(model.obj,newdata=predictors),levels=c(0,1)),observed)
    # } else if(model.string=="GLM"){
    #   # observed <- as.vector(y)
    #   # CM <- caret::confusionMatrix(
    #   #   as.factor(predict(model.obj,
    #   #                     as.matrix(predictors[, -which(colnames(predictors) ==
    #   #                                                     "LPS.Added")]),type="class",s='lambda.min')),
    #   #   as.factor(observed))
    #   observed <- as.factor(y)
    #   CM <- caret::confusionMatrix(as.factor(ifelse(predict(model.obj,newdata = predictors,type="response")>.5,0,1)),
    #                         observed)
    # } 
  error.vec <- c(CM$overall,CM$byClass)
  return(error.vec)
}
## GLM Elastic Net
p <- predict(cv_glm,newx = as.matrix(train_data_scaled[,-c(1)]),s=.44,type="response") # Probabilities(of 1)
as.factor(predict(cv_glm,newx = as.matrix(train_data_scaled[,-c(1)]),s=.44,type="class")) # class
## NN
predict(rfe_nn$fit,newdata=train_data_scaled,type="prob")$`1` # Probabilities
predict(rfe_nn$fit,newdata=train_data_scaled,type="raw") # class
## SVM
library(kernlab)
# vars_incl <- colnames(rfe_svm$fit$trainingData)[1:(length(colnames(rfe_svm$fit$trainingData))-1)]
rebuild_svm <- ksvm(LPS.Added~.,data=train_data_scaled[,c("LPS.Added",vars_incl)],type="C-svc",kernel="vanilladot",prob.model=TRUE,C=.1)
predict(rebuild_svm,newdata=train_data_scaled,type="probabilities")[,2] #Prob
predict(rebuild_svm,newdata=train_data_scaled,type="response") #class

### RF
predict(rfe_rf$fit,newdata=train_data_scaled,type="response")#class
predict(rfe_rf$fit,newdata=train_data_scaled,type="prob")[,2]#Prob

### NB
predict(rfe_nb$fit,newdata=train_data_scaled,type="response")$class#class
predict(rfe_nb$fit,newdata=train_data_scaled,type="prob")$posterior[,2]#Prob

generate.errors.ML <- function(model.string,model.obj,predictors){
  y <- predictors$LPS.Added
  if(model.string %in% c("RF")){
    observed <- as.factor(y)
    class_pred <- as.factor(predict(model.obj,predictors,type="response"))
    prob_pred <- predict(model.obj,predictors,type="prob")[,2]
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(model.string%in%c("SVM")){
    observed <- as.factor(y)  
    class_pred <- as.factor(predict(model.obj,predictors,type="response"))
    prob_pred <- predict(model.obj,predictors,type="probabilities")[,2]
    colnames(prob_pred) <- NULL
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(grepl("GLM_EN",model.string)){
    # predict(cv_glm,newx = as.matrix(train_data_scaled[,-c(1)]),s=.44,type="response") # Probabilities(of 1)
    # predict(cv_glm,newx = as.matrix(train_data_scaled[,-c(1)]),s=.44,type="class") # class
    observed <- as.factor(y)
    class_pred <- as.factor(predict(model.obj,as.matrix(predictors[-c(1)]),type="class",s=.44))
    prob_pred <- predict(model.obj,as.matrix(predictors[-c(1)]),type="response",s=.44)
    CM <- caret::confusionMatrix(class_pred,observed)
  } else if(grepl("NB",model.string)){
    observed <- as.factor(y)
    class_pred <- as.factor(predict(model.obj,predictors,type="response")$class)
    prob_pred <- predict(model.obj,predictors,type="prob")$posterior[,2]
    CM <- caret::confusionMatrix(class_pred,observed)
  }else if(grepl("NN",model.string)){
    observed <- as.factor(y)  
    class_pred <- as.factor(predict(model.obj,predictors,type="raw"))
    prob_pred <- predict(model.obj,predictors,type="prob")[,2]
    CM <- caret::confusionMatrix(class_pred,observed)
  }
  error.vec <- c(CM$overall,CM$byClass)
  return(list(error.vec,data.frame(prob_1=prob_pred,prob_0=1-prob_pred,class_pred=class_pred,observed=observed) %>% rename(prob_1=1,prob_0=2)))
}

generate.errors.ML("GLM_EN",cv_glm,train_data_scaled)

errors.all.data.conf <- function(model.string,model.obj){
  error.df <- generate.errors.ML(model.string,model.obj,train_data_scaled)[[2]] %>% 
    mutate(model=model.string) %>% 
    mutate(dataset="train") %>% 
    bind_rows(generate.errors.ML(model.string,model.obj,test_data_scaled)[[2]] %>% 
    mutate(model=model.string) %>% 
    mutate(dataset="test"))%>% 
    bind_rows(generate.errors.ML(model.string,model.obj,MJ.LPS_data_scaled)[[2]] %>% 
    mutate(model=model.string) %>% 
    mutate(dataset="MJ"))
  return(error.df)
}
err_conf <- bind_rows(errors.all.data.conf("RF",rfe_rf$fit),
                     errors.all.data.conf("NB",rfe_nb$fit),
                     errors.all.data.conf("SVM",rebuild_svm),
                     errors.all.data.conf("NN",rfe_nn$fit),
                     errors.all.data.conf("GLM_EN",cv_glm)) %>% 
  mutate(prob_observed=ifelse(observed==0,prob_0,prob_1))

ggplot(err_conf ,aes(x=model,y=prob_observed,color=dataset,fill=dataset))+
  geom_violin(scale="width",alpha=.3)#+
  # facet_grid(.~model)

ggplot(err_conf ,aes(x=prob_observed,color=dataset))+
  geom_density()+
  facet_wrap(~model,scales = "free")

ggplot(err_conf,aes(x=prob_observed,color=dataset,fill=observed))+
  geom_density()+
  facet_wrap(.~model,scales = "free")
errors.all.data <- function(model.string,model.obj){
  error.df <- data.frame(
    cbind(
      train=generate.errors.ML(model.string,model.obj,train_data_scaled)[[1]],
      test=generate.errors.ML(model.string,model.obj,test_data_scaled)[[1]],
      mj_LPS=generate.errors.ML(model.string,model.obj,MJ.LPS_data_scaled)[[1]]
      )
  ) %>% 
    mutate(model=model.string) %>% 
    select(model,train:mj_LPS) %>% 
    rownames_to_column(var="metric")
}
RF.errors <- errors.all.data("RF",rfe_rf$fit) 
NB.errors <- errors.all.data("NB",rfe_nb$fit) 
SVM.errors <- errors.all.data("SVM",rebuild_svm) 
NN.errors <- errors.all.data("NN",rfe_nn$fit)
GLM.errors <- errors.all.data("GLM_EN",cv_glm)
print(xtable(rbind(RF.errors[1,-6],
      NB.errors[1,-6],
      SVM.errors[1,-6],
      NN.errors[1,-6],
      GLM.errors[1,-6]
      ),digits=4),include.rownames=FALSE)
```

```{r ROC}
library(pROC)
predicted_rf <-  predict(rfe_rf$fit,test_data_scaled,type = "prob")[,2]
roc_obj_rf <- roc(test_data_scaled$LPS.Added,predicted_rf)
auc_rf <- round(auc(test_data_scaled$LPS.Added,predicted_rf),4)

#create ROC plot
ggroc(roc_obj_rf, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc_rf, ')'))
roc_obj_nn <- roc(as.numeric(test_data_scaled$LPS.Added), predict(rfe_nn$fit,test_data_scaled,type = "prob")[,2])
ggroc(roc_obj_nn)


roc_obj_glm <- roc(as.numeric(test_data_scaled$LPS.Added), predict(cv_glm,as.matrix(test_data_scaled[,-1]),type = "response")[,2])
ggroc(roc_obj_glm)

roc_obj_svm <- roc(as.numeric(test_data_scaled$LPS.Added), predict(rebuild_svm,test_data_scaled,type = "prob")[,2])
ggroc(roc_obj_svm)


roc_obj_nb <- roc(as.numeric(test_data_scaled$LPS.Added), predict(rfe_nb$fit,test_data_scaled,type = "prob")$posterior[,2])
ggroc(roc_obj_nb)

roc_auc_fxn <- function(model.string,model.obj){
  predictions<- bind_rows(generate.errors.ML(model.string,model.obj,train_data_scaled)[[2]] %>%
    mutate(datatype = "train")%>% 
    mutate(auc_metric = as.numeric(auc(observed,prob_1))),
    generate.errors.ML(model.string,model.obj,test_data_scaled)[[2]] %>%
    mutate(datatype = "test")%>% 
    mutate(auc_metric = as.numeric(auc(observed,prob_1))),
    generate.errors.ML(model.string,model.obj,MJ.LPS_data_scaled)[[2]] %>%
    mutate(datatype = "MJ")%>% 
    mutate(auc_metric = as.numeric(auc(observed,prob_1)))
  ) %>% 
    mutate(model= model.string) %>% 
    select(datatype, auc_metric,model) %>% distinct()
  return(predictions)
}

RF.auc <- roc_auc_fxn("RF",rfe_rf$fit) 
NB.auc <- roc_auc_fxn("NB",rfe_nb$fit) 
SVM.auc <- roc_auc_fxn("SVM",rebuild_svm) 
NN.auc <- roc_auc_fxn("NN",rfe_nn$fit)
GLM.auc <- roc_auc_fxn("GLM_EN",cv_glm)
print(xtable(rbind(RF.auc,
      NB.auc,
      SVM.auc,
      NN.auc,
      GLM.auc
      ),digits=4),include.rownames=FALSE)

```
